---
title: "Project codes"
author: "Latera Tesfaye Olana"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
  word_document: default
header-includes: 
 \usepackage{float}
 \usepackage{booktabs}
 \usepackage{longtable}
 \usepackage{array}
 \usepackage{multirow}
 \usepackage[table]{xcolor}
 \usepackage{wrapfig}
 \usepackage{float}
 \floatplacement{figure}{H}
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(knitr)
library(gridExtra)
library(ggpubr)
library(ggfx)
library(prioritizr)
library(cowplot)
library(SUMMER)
library(forcats)
library(rdhs)
library(INLA)
library(survey)
library(readr)
library(viridis)
library(surveyPrev)
library(patchwork)
library(tidyr)
library(patchwork)
library(kableExtra)
library(sf)
library(tidyverse)
library(ggrepel)
library(ggstance)
library(scales)
library(igraph)   
library(dplyr)    
library(terra)
library(exactextractr)
library(INLA)
library(spdep)
library(ggtext)
library(forcats)
library(survey)
library(viridis)
library(ggnewscale)
library(ggspatial)
library(patchwork)
library(purrr)
library(msm) 
library(showtext)
library(grid)
library(RColorBrewer)



knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(
  
  fig.pos = 'H', # Places figures at top or here
  out.width = '100%', dpi = 300, # Figure resolution and size
  fig.env="figure"
) # Latex figure environment
#library(devtools)
#install_github("richardli/surveyPrev")
options(knitr.table.format = "latex") # For kable tables to write LaTeX table directly
```











```{r echo=TRUE, message=FALSE}
#data
data <- read_csv("C:/Users/latera/OneDrive - UW/Desktop/Proposal/Thesis/Codes/Data/416582.csv")
 data_2019 <- data 
 data_2019 <- data_2019 %>% filter(
   age_year == 1
 )

data_2019 <- data_2019 %>% rename(
  starta1 = strata
)
data_2019 <- data_2019 %>%
  dplyr::rename(
    # cluster = V001,
    # householdID = V002,
    region = admin_1,
    # weight = V005,
    strata = urban,
    value = mcv1
  ) 
data_2019 <- data_2019 %>% dplyr::select(
  c(cluster,householdID,region,weight,strata,value)
)
data_2019 <- data_2019%>% dplyr::rename(
  v024 = region
)
data_2019$strata <- ifelse(data_2019$strata == 1, "urban", "rural")
data_2019 <- as.data.frame(data_2019)
data_2019$strata <- as.factor(data_2019$strata)
data_2019$v024 <- as.factor(data_2019$v024)

```

Assessing this data tells us, `r nrow(data)` observations, `r length(unique(data$cluster))`, and `r length(unique(data$householdID))`. 

The following code reads the geographic information. 

```{r echo=TRUE, message=FALSE, results = 'hide'}

geo <- st_read("ETGE81FL/ETGE81FL.shp")

# geo <- geo %>%
#   mutate(ADM1NAME = case_when(
#     ADM1NAME == "Afar Region" ~ "Afar",
#     ADM1NAME == "Amhara Region" ~ "Amhara",
#     ADM1NAME == "Oromiya Region" ~ "Oromia",
#     ADM1NAME == "Tigray Region" ~ "Tigray",
#     ADM1NAME == "Benishangul Gumuz" ~ "Benishangul Gumz",
#     ADM1NAME == "SNNPR" ~ "SNNP",
#     ADM1NAME == "Gambella" ~ "Gambela",
#     TRUE ~ ADM1NAME  # Keeps all other values as they are
#   ))
```



```{r echo=TRUE, message=FALSE, fig.cap="The cluster distribution of DHS 2016 in Ethiopia"}
ggplot(data = geo) +
  geom_sf(fill = NA, 
          color = "steelblue", 
          size = 2, 
          shape = 1) +  
  theme_minimal() +
  labs(title = "",
       caption = "",
       x = "Longitude",
       y = "Latitude") +
   theme_minimal() +
  theme(legend.position = "bottom",
        legend.background = 
          element_rect(fill = "transparent"),
        legend.box.background = 
          element_rect(color = "transparent"),
        plot.title = element_text(color = "black", 
                                  size = 15, 
                                  face = "bold"),
        plot.subtitle = element_text(color = "black", 
                                     size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank()) + 
  coord_sf()


```



```{r}
merged_data_new <- left_join(data_2019, geo, by = c("cluster" = "DHSCLUST"))
```


```{r echo=TRUE, message=FALSE, results = 'hide'}
data_ethiopia_1 <- 
  st_read("ETH_wereda_Bofed_CSA_2021/ETH_wereda_Bofed_CSA_2021.shp")


drc1_sf <- st_as_sf(merged_data_new, coords = c("LONGNUM", "LATNUM"), crs = st_crs(data_ethiopia_1), remove = FALSE)

merged_data_sf <- st_join(data_ethiopia_1, drc1_sf, join = st_contains)

# st_write(merged_data_sf, "shapefile_containing_individual_and _cluster_level_data.shp")

```

In the following code we will create admin-1 and admin-2 datasets. 


```{r echo=TRUE, message=FALSE}
admin1_boundaries <- data_ethiopia_1 %>%
  group_by(ADM1_EN) %>%
  dplyr::summarize(geometry = st_union(geometry))

admin2_boundaries <- data_ethiopia_1 %>%
  group_by(ADM2_EN, ADM1_EN) %>%
  dplyr::summarize(geometry = st_union(geometry))
```

```{r}
df <- st_drop_geometry(merged_data_sf)
df2 <- df %>%
  filter(!is.na(value))
adm2_summary <- df2 %>%
  group_by(ADM2_EN) %>%
  summarise(
    distinct_vals = n_distinct(value),
    # mark TRUE if only one distinct value (either all 0s or all 1s)
    homogeneous = (distinct_vals == 1),
    .groups = "drop"
  )
total_units       <- nrow(adm3_summary)
homogeneous_units <- sum(adm3_summary$homogeneous)
pct_homogeneous   <- homogeneous_units / total_units * 100

pct_homogeneous
homogeneous_units
total_units
```


```{r}
cluster_agg <- merged_data_sf %>%
  st_drop_geometry() %>%
  group_by(cluster) %>%
  summarise(n_sample = sum(!is.na(value)))
cluster_points <- drc1_sf %>%
  group_by(cluster) %>%
  slice(1) %>%
  ungroup()
cluster_counts <- left_join(cluster_points, cluster_agg, by = "cluster")
adm2_boundaries <- data_ethiopia_1 %>%
  group_by(ADM2_EN) %>%
  summarize(geometry = st_union(geometry)) %>%
  st_as_sf()

adm1_boundaries <- data_ethiopia_1 %>%
  group_by(ADM1_EN) %>%
  summarize(geometry = st_union(geometry)) %>%
  st_as_sf()
adm3_sample_data <- cluster_counts %>%
  st_join(
    data_ethiopia_1 %>% select(ADM3_EN),  
    join = st_within
  ) %>%
  st_drop_geometry() %>%
  group_by(ADM3_EN) %>%
  summarise(total_samples = sum(n_sample, na.rm = TRUE)) %>%
  right_join(data_ethiopia_1, by = "ADM3_EN") %>%  
  st_as_sf()



adm2_sample_data <- cluster_counts %>%
  st_join(
    adm2_boundaries %>% select(ADM2_EN),  
    join = st_within
  ) %>%
  st_drop_geometry() %>%
  group_by(ADM2_EN) %>%
  summarise(total_samples = sum(n_sample, na.rm = TRUE)) %>%
  right_join(adm2_boundaries, by = "ADM2_EN") %>%  
  st_as_sf()
p1 <- ggplot() +
  geom_sf(
    data = adm3_sample_data,
    aes(fill = total_samples),
    color = "black", 
    linewidth = 0.15
  ) +
  
  #ADM2 boundaries (bold black lines)
  
  
  geom_sf(
  data = adm1_boundaries,
  fill = NA,
  color = "black",
  alpha = 0.7,
  linewidth = 0.7
) +
geom_sf(
  data = adm1_boundaries,
  color = "gray34",
  linewidth = 1.1,
  alpha = 0.3
) +
  geom_sf(
    data = cluster_counts,
    color = "red",
    # shape=21,
    alpha=0.65,
    size = 1.5
  ) +
  scale_fill_viridis_c(
    option = "mako",
    name = "Samples: ",
    direction = -1,
    na.value = "#f8f9fa", 
    breaks = scales::pretty_breaks(n = 6),
    guide = guide_colorbar(
      barwidth = 12,
      barheight = 0.8,
      title.position = "top"
    )
  ) +
  labs(
    title = "",
    subtitle = "",
    caption = ""
  ) +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(
      size = 22,
      face = "bold",
      hjust = 0.5,
      margin = margin(b = 8),
      color = "#2d3748"
    ),
    plot.subtitle = element_text(
      size = 16,
      hjust = 0.5,
      margin = margin(b = 20),
      color = "#4a5568"
    ),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = unit(c(1, 1, 1.5, 1), "cm")
  )
p2 <- ggplot() +
  geom_sf(
    data = adm2_sample_data,
    aes(fill = total_samples),
    color = "black",  # Thin white boundaries between ADM3 units
    linewidth = 0.15
  ) +
  geom_sf(
  data = adm1_boundaries,
  fill = NA,
  color = "black",
  alpha = 0.7,
  linewidth = 0.7
) +
geom_sf(
  data = adm1_boundaries,
  color = "gray34",
  linewidth = 1.1,
  alpha = 0.3
) +
  geom_sf(
    data = cluster_counts,
    color = "red",
    # shape=21,
    alpha=0.65,
    size = 1.5
  ) +
  scale_fill_viridis_c(
    option = "mako",
    name = "Samples: ",
    direction = -1,
    na.value = "#f8f9fa",  # Light gray for districts with no samples
    breaks = scales::pretty_breaks(n = 6),
    guide = guide_colorbar(
      barwidth = 12,
      barheight = 0.8,
      title.position = "top"
    )
  ) +
  labs(
    title = "",
    subtitle = "",
    caption = ""
  ) +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(
      size = 22,
      face = "bold",
      hjust = 0.5,
      margin = margin(b = 8),
      color = "#2d3748"
    ),
    plot.subtitle = element_text(
      size = 16,
      hjust = 0.5,
      margin = margin(b = 20),
      color = "#4a5568"
    ),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = unit(c(1, 1, 1.5, 1), "cm")
  )


final_plot <- p1 + p2 +
  plot_layout(ncol = 2) 
print(final_plot)


 ggsave("map1.png", width = 14, height = 10, bg = "white")



final_plot

```




```{r}
merged_non_na <- merged_data_sf %>% filter(
  !is.na(value)
)

design_based <- svydesign(
  ids = ~cluster + householdID, 
  weights = ~weight, 
  starta = ~strata,
  data = merged_non_na
)

design_based_estimates <- svyby(~value, ~ADM3_EN, design_based, svymean)

design_based_estimates_shp <- left_join(data_ethiopia_1, design_based_estimates, 
                                        by="ADM3_EN")

# st_write(design_based_estimates_shp, "result/direct_estimate_admin3.shp")
```


```{r}
design_based_estimates_shp$value <- design_based_estimates_shp$value * 100
design_based_estimates_shp$se <- design_based_estimates_shp$se * 100
plot2 <- ggplot() +
  geom_sf(data = subset(design_based_estimates_shp_admin2, !is.na(value) & se > 0.0005), 
          aes(fill = value), color = "white", lwd = 0.05) +
  scale_fill_viridis(
    name = "Measles Coverage (%)",
    option = "mako",  
    begin = 0.15,
    end = 0.85,
    direction = -1,
    na.value = "transparent",
    limits = c(0, 100),
    guide = guide_colorbar(
      direction = "horizontal",
      barheight = unit(2, "mm"),
      barwidth = unit(80, "mm"),
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "gray30",
      ticks.colour = "gray30")
  ) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = subset(design_based_estimates_shp_admin2, se < 0.0005),
          fill = "#E63946", color = "white", lwd = 0.05) +  # Scientific red
  geom_sf(data = subset(design_based_estimates_shp_admin2, is.na(value)),
          fill = "#495057", color = "white", lwd = 0.05) +  
  geom_sf(
  data = adm1_boundaries,
  fill = NA,
  color = "black",
  alpha = 0.7,
  linewidth = 1.1
) +
  theme_void() +
  theme(
    # plot.background = element_rect(fill = "#F8F9FA", color = NA),
    # panel.background = element_rect(fill = "#F8F9FA", color = NA),
    legend.position = "bottom",
    legend.title = element_text(
      size = 20,
      face = "bold",
      margin = margin(b = 3)),
    legend.text = element_text(size = 17),
    plot.title = element_text(
      face = "bold",
      size = 20,
      hjust = 0.5,
      color = "#2C3E50",
      margin = margin(b = 8)),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      color = "#7F8C8D",
      margin = margin(b = 15)),
    plot.caption = element_text(
      size = 9,
      color = "#95A5A6",
      hjust = 0.98,
      margin = margin(t = 10))
  ) +
  labs(
    title = "",
    subtitle = "",
    caption = ""
  ) +
  annotate("text", 
           x = Inf, y = Inf,
           label = c("Missing Data", "Zero Standard Error"),
           color = c("#2F4F4F", "#DC143C"),
           hjust = c(1.1, 1.1),
           vjust = c(1.5, 3),
           size = 8,
           fontface = "bold")


ggsave("measles_coverage_final_admin1_sc.png",
       width = 13, height = 9, dpi = 400, bg = "white")
```


```{r}

design_based <- svydesign(
  ids = ~cluster + householdID, 
  weights = ~weight, 
  starta = ~strata,
  data = merged_non_na
)

design_based_estimates_admin2 <- svyby(~value, ~ADM2_EN, design_based, svymean)

design_based_estimates_shp_admin2 <- left_join(adm2_boundaries, design_based_estimates_admin2, 
                                        by="ADM2_EN")

st_write(design_based_estimates_shp_admin2, "result/direct_estimate_admin2.shp")


```


```{r}
design_based_estimates_shp_admin2$value <- design_based_estimates_shp_admin2$value * 100
design_based_estimates_shp_admin2$se <- design_based_estimates_shp_admin2$se * 100

#
# plot2 <- ggplot() +
#   # Main coverage data
#   geom_sf(data = subset(design_based_estimates_shp_admin2, !is.na(value) & se != 0), 
#           aes(fill = value), color = "white", lwd = 0.05) +
#    scico::scale_fill_scico(
#         palette = "navia",             #viridis
#         direction = -1,               
#         begin = 0.1,                  
#          name = "DPT1",
#         end = 0.9,
#         guide = guide_colorbar(
#             direction = "horizontal",
#             barwidth = unit(8, "mm"),
#             barheight = unit(100, "mm"),
#             title.position = "top",
#             frame.colour = "#404040",
#             ticks = FALSE,
#             title.theme = element_text(
#                 size = 10,
#                 color = "#404040",
#                 face = "italic"
#             )
#         )
#     ) +
#   geom_sf(data = subset(design_based_estimates_shp_admin2, se == 0),
#           fill = "#E63946", color = "white", lwd = 0.05) +  
#   geom_sf(data = subset(design_based_estimates_shp_admin2, is.na(value)),
#           fill = "#495057", color = "white", lwd = 0.05) +  
#   
#   #professional north arrow
#   
#    geom_sf(
#   data = adm1_boundaries,
#   fill = NA,
#   color = "black",
#   alpha = 0.7,
#   linewidth = 1.1
# ) +
#
#   # annotation_scale(
#   #   location = "bl",
#   #   style = "ticks",
#   #   line_width = 0.4,
#   #   text_cex = 0.9,
#   #   height = unit(0.2, "cm"),
#   #   pad_x = unit(0.2, "cm"),
#   #   pad_y = unit(0.2, "cm")
#   # ) +
#   
#   theme_void() +
#   theme(
#     # plot.background = element_rect(fill = "#F8F9FA", color = NA),
#     # panel.background = element_rect(fill = "#F8F9FA", color = NA),
#     legend.position = c(-0.01, 0.5),
#         legend.title = element_text(
#             size = 11,
#             color = "#404040",
#             margin = margin(b = 4)
#         ),
#     legend.text = element_text(size = 17),
#   
#     plot.title = element_text(
#       face = "bold",
#       size = 20,
#       hjust = 0.5,
#       color = "#2C3E50",
#       margin = margin(b = 8)),
#     plot.subtitle = element_text(
#       size = 12,
#       hjust = 0.5,
#       color = "#7F8C8D",
#       margin = margin(b = 15)),
#     plot.caption = element_text(
#       size = 9,
#       color = "#95A5A6",
#       hjust = 0.98,
#       margin = margin(t = 10))
#   ) +
#   
#   labs(
#     title = "",
#     subtitle = "",
#     caption = ""
#   ) +
#   
#   annotate("text", 
#            x = Inf, y = Inf,
#            label = c("Missing Data", "Zero Standard Error"),
#            color = c("#2F4F4F", "#DC143C"),
#            hjust = c(1.1, 1.1),
#            vjust = c(1.5, 3),
#            size = 8,
#            fontface = "bold")

final_plot <- plot1 + plot2 +
  plot_layout(ncol = 2) 
print(final_plot)

ggsave("corrected_both_admin.png", 
       width = 16, height = 12, dpi = 400, bg = "white")
```

```{r}
design_based <- svydesign(
  ids = ~cluster + householdID, 
  weights = ~weight, 
  strata = ~strata,
  data = merged_non_na
)
design_based_estimates_admin1 <- svyby(~value, ~ADM1_EN, design_based,vartype  = c("se", "ci"), svymean)

design_based_estimates_shp_admin1 <- left_join(adm1_boundaries, design_based_estimates_admin1, 
                                        by="ADM1_EN")

design_based_estimates_shp_admin1$value <- design_based_estimates_shp_admin1$value * 100
design_based_estimates_shp_admin1$se <- design_based_estimates_shp_admin1$se * 100
design_based_estimates_shp_admin1

names(design_based_estimates_shp_admin1)
plot_data <- design_based_estimates_shp_admin1 |>
  mutate(
    ADM1_EN = fct_reorder(ADM1_EN, value),
    ci_lower = value - 1.96 * se,
    ci_upper = value + 1.96 * se,
    above_90 = ifelse(value > 90, "Above 90%", "Below 90%")
  )
plot_data <- st_drop_geometry(plot_data)

plot_data$ci_upper <- pmin(plot_data$ci_upper, 100)
ggplot(plot_data, aes(x = value, y = ADM1_EN, color = above_90)) +
  # ci
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper), 
                 height = 0.2, size = 0.8, alpha = 0.7) +
  # pe
  geom_point(size = 3, shape = 19) +
  geom_vline(xintercept = 90, linetype = "dashed", color = "#e63946", size = 0.8) +
  geom_text(aes(label = scales::percent(value/100, accuracy = 1)), 
            hjust = -0.4, vjust=-0.1, size = 3, family = "roboto", show.legend = FALSE) +
  scale_color_manual(values = c("Above 90%" = "#2a9d8f", "Below 90%" = "#264653")) +
  scale_x_continuous(
  breaks = c(seq(0, 100, by = 20), 90),
  labels = scales::percent_format(scale = 1)
) +
  labs(title = "",
       subtitle = "",
       x = "Estimated Coverage (%)",
       y = "",
       caption = "",
       color = "Coverage Status") +
  theme_minimal(base_size = 12, base_family = "roboto") +
  theme(
    plot.title = element_markdown(hjust = 0.5, margin = margin(b = 10)),
    plot.subtitle = element_text(hjust = 0.5, color = "gray40", size = 14),
    plot.caption = element_text(color = "gray50", hjust = 0.95, margin = margin(t = 15)),
    panel.grid.major.y = element_line(color = "gray90", size = 0.2),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "gray30"),
    axis.title.x = element_text(margin = margin(t = 10), color = "gray30"),
    legend.position = "top",
    legend.justification = "right",
    legend.title = element_text(size = 10),
    plot.background = element_rect(fill = "#ffffff", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = unit(c(1.5, 2, 1, 2), "cm")
  ) 
```

```{r}
ggsave("admin1_coverage.png", 
       width = 8, height = 5, dpi = 400, bg = "white")
```


```{r}

merged_data_sf <- st_join(admin2_boundaries, drc1_sf, join = st_contains)

merged_data_sf_id <- merged_data_sf %>%
  mutate(ADM2_ID = as.integer(as.factor(ADM2_EN)))

adm2_boundaries_id <- merged_data_sf_id %>%
  group_by(ADM2_EN, ADM2_ID) %>%
  summarize(geometry = st_union(geometry)) %>%
  st_as_sf()

adm2_boundaries_id$ADM2_ID_iid <- adm2_boundaries_id$ADM2_ID

binom_df <- merged_data_sf_id %>% 
  group_by(cluster,ADM2_ID) %>% 
  summarise(y = sum(value, na.rm = T), 
            N = length(which(!(is.na(value)))),
            ADM2_EN = ADM2_EN %>% nth(1), 
            ADM2_ID = ADM2_ID %>% nth(1),
            urban = as.numeric(nth(strata, 1) == "urban"), 
            rural = as.numeric(nth(strata, 1) == "rural")) %>%
  as.data.frame()

binom_df$ADM2_ID_iid <- binom_df$ADM2_ID


binom_sf <- st_as_sf(binom_df)

nb_adm1 <- poly2nb(adm2_boundaries_id, row.names = adm2_boundaries_id$ADM2_ID)
adj_matrix <- nb2mat(nb_adm1, style = "B", zero.policy = TRUE)
adj_list_adm1 <- inla.read.graph(adj_matrix)


formula <- y ~ urban + rural - 1 +
  f(ADM2_ID, model = "bym2", graph = adj_list_adm1) 

binom_df$y[binom_df$N == 0] <- NA

result <- inla(formula = formula, 
               Ntrials = N,
               data = binom_df,
               family = "betabinomial",
               control.predictor = list(compute = TRUE),
               control.compute = list(return.marginals = TRUE,
                                      config = TRUE,
                                      return.marginals.predictor = TRUE))


# urban_prop is calculated from .tif file below


files <- c(
  "GHS_SMOD_E2030_GLOBE_R2023A_54009_1000_V2_0_R9_C23.tif",
  "GHS_SMOD_E2030_GLOBE_R2023A_54009_1000_V2_0_R9_C22.tif",
  "GHS_SMOD_E2030_GLOBE_R2023A_54009_1000_V2_0_R8_C23.tif",
  "GHS_SMOD_E2030_GLOBE_R2023A_54009_1000_V2_0_R8_C22.tif"
)

# SpatRasters
smod_list <- lapply(files, rast)

# Mosaic them into a single SpatRaster
# do.call(...) helps combine them
smod_mosaic <- do.call(mosaic, smod_list)
smod_mosaic

eth_adm2 <- adm2_boundaries_id
eth_adm2_moll <- st_transform(eth_adm2, crs(smod_mosaic))
eth_vect_moll <- vect(eth_adm2_moll)
smod_crop <- crop(smod_mosaic, eth_vect_moll)
smod_ethiopia <- mask(smod_crop, eth_vect_moll)
plot(smod_ethiopia)
reclass_mat <- matrix(c(
  10, 0,  # Water -> 0
  11, 0,  # Very low density rural -> 0
  12, 0,  # Low density rural -> 0
  13, 0,  # Rural cluster -> 0
  21, 1,  # Suburban/peri-urban -> 1
  22, 1,  # Semi-dense cluster -> 1
  23, 1,  # Dense urban cluster -> 1
  30, 1   # Urban centre -> 1
), ncol=2, byrow=TRUE)
#reclassify
urban_mask <- classify(smod_ethiopia, reclass_mat, include.lowest=FALSE, right=NA)

eth_adm2$urban_prop <- exact_extract(
  urban_mask, 
  eth_adm2,
  fun = function(values, coverage_fraction) {
    weighted.mean(values, coverage_fraction, na.rm = TRUE)
  }
)
urban_prop <- data.frame(
  urban_proportion <- eth_adm2$urban_prop,
  ADM2_ID <- eth_adm2$ADM2_ID,
  ADM2_EN <- eth_adm2$ADM2_EN
)

# ggplot(eth_adm2) +
#   geom_sf(aes(fill = urban_prop), color = "white", size=0.2) +
#   scale_fill_viridis_c(option="C", name="Urban Fraction") +
#   labs(title="Urban Proportion by ADM2 (GHSL SMOD)") +
#   theme_minimal()

# urban_prop <- binom_df %>%
#   group_by(ADM2_EN, ADM2_ID) %>%
#   summarize(
#     total_observations = n(),
#     urban_count = sum(urban == 1, na.rm = TRUE),
#     urban_proportion = mean(urban == 1, na.rm = TRUE)
#   ) %>%
#   ungroup()

urban_prop$urban_proportion[is.nan(urban_prop$urban_proportion)] <- 0
nsamp <- 1000
samp <- inla.posterior.sample(n = nsamp, result = result)
region_idx <- which(rownames(samp[[1]]$latent) %>% 
                      str_detect("ADM2"))[1:length(unique(binom_df$ADM2_ID))]

urban_idx <- which(rownames(samp[[1]]$latent) %>% str_detect("urban"))
rural_idx <- which(rownames(samp[[1]]$latent) %>% str_detect("rural"))

region_mat <- matrix(0, nrow = length(region_idx), ncol = nsamp)
urban_mat <- matrix(0, nrow = 1, ncol = nsamp)
rural_mat <- matrix(0, nrow = 1, ncol = nsamp)
# matrix with posterior samples
for (i in 1:nsamp) {
  region_mat[,i] <- samp[[i]]$latent[region_idx]
  urban_mat[,i] <- samp[[i]]$latent[urban_idx]
  rural_mat[,i] <- samp[[i]]$latent[rural_idx]
}
samp_mat_urban <- region_mat + urban_mat[rep(1, nrow(region_mat)),]
samp_mat_rural <- region_mat + rural_mat[rep(1, nrow(region_mat)),]
expit_samp_mat_urban <- expit(samp_mat_urban)
expit_samp_mat_rural <- expit(samp_mat_rural)
expit_samp_mat <- (urban_prop$urban_proportion * expit_samp_mat_urban + (1-urban_prop$urban_proportion) * expit_samp_mat_rural)

samp_betabinom <- expit_samp_mat

bino <- data.frame(
  ADM2_ID = sort(unique(adm2_boundaries_id$ADM2_ID)),
  ADM2_EN = sort(unique(adm2_boundaries_id$ADM2_EN)),
  med_beta_bino = apply(samp_betabinom, 1, median),
  lower_beta_bino = apply(samp_betabinom, 1, quantile, 0.025),
  upper_beta_bino = apply(samp_betabinom, 1, quantile, 0.975),
  sd_beta_bino = apply(samp_betabinom, 1, sd)
)

#data_ethiopia_1 is a shape file for admin-2
plot_plot <- left_join(adm2_boundaries_id, bino, by="ADM2_ID")

plot_plot$med_beta_bino <- plot_plot$med_beta_bino * 100
plot_plot$sd_beta_bino <- plot_plot$sd_beta_bino * 100

st_write(plot_plot, "result/unit_level_estimate_admin2_urban_new_new.shp")
ggplot() +
  #coverage data
  geom_sf(data = subset(plot_plot, !is.na(med_beta_bino) & sd_beta_bino != 0), 
          aes(fill = med_beta_bino), color = "white", lwd = 0.05) +
  scale_fill_viridis(
    name = "Measles Coverage (%)",
    option = "mako", 
    direction = -1,
    begin = 0.15,
    end = 0.85,
    na.value = "transparent",
    limits = c(0, 100),
    guide = guide_colorbar(
      direction = "horizontal",
      barheight = unit(2, "mm"),
      barwidth = unit(80, "mm"),
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "gray30",
      ticks.colour = "gray30")
  ) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = subset(plot_plot, sd_beta_bino == 0),
          fill = "#E63946", color = "white", lwd = 0.05) +  
  geom_sf(data = subset(plot_plot, is.na(med_beta_bino)),
          fill = "#495057", color = "white", lwd = 0.05) +  
  geom_sf(
  data = adm1_boundaries,
  fill = NA,
  color = "black",
  alpha = 0.7,
  linewidth = 1.1
) +
  
#this NEW layer right after existing admin1 layer

  # Scale bar with refined style
  # annotation_scale(
  #   location = "bl",
  #   style = "ticks",
  #   line_width = 0.4,
  #   text_cex = 0.9,
  #   height = unit(0.2, "cm"),
  #   pad_x = unit(0.2, "cm"),
  #   pad_y = unit(0.2, "cm")
  # ) +
  theme_void() +
  theme(
    # plot.background = element_rect(fill = "#F8F9FA", color = NA),
    # panel.background = element_rect(fill = "#F8F9FA", color = NA),
    legend.position = "bottom",
    legend.title = element_text(
      size = 20,
      face = "bold",
      margin = margin(b = 3)),
    legend.text = element_text(size = 17),
  
    plot.title = element_text(
      face = "bold",
      size = 20,
      hjust = 0.5,
      color = "#2C3E50",
      margin = margin(b = 8)),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      color = "#7F8C8D",
      margin = margin(b = 15)),
    plot.caption = element_text(
      size = 9,
      color = "#95A5A6",
      hjust = 0.98,
      margin = margin(t = 10))
  ) +
  labs(
    title = "",
    subtitle = "",
    caption = ""
  ) 


#here we have used urban prop form GSHL
ggsave("measles_coverage_final_admin2_cluster_level_new.png", 
       width = 16, height = 12, dpi = 400, bg = "white")
```


```{r}
ggplot(eth_adm3) +
  geom_sf(aes(fill = urban_prop), color = "gray70", size = 0.2) +
  scale_fill_distiller(
    palette   = "PuBu",   
    direction = 1,
    name      = "Urban Fraction",
    guide = guide_colorbar(
      barwidth  = 15,
      barheight = 1
    )
  ) +
  labs(
    title    = "",
    subtitle = "",
    caption  = ""
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title       = element_text(face = "bold"),
    plot.subtitle    = element_text(margin = margin(b = 10)),
    axis.title       = element_blank(),
    axis.text        = element_blank(),
    panel.grid       = element_blank(),
    legend.position  = "bottom"
  )

ggsave("urban_map_admin3.png", 
       width = 13, height = 8, dpi = 400, bg = "white")
```


```{r}


# ------------------------------------------------------------------------------
#Convert polygons to centroids
# ------------------------------------------------------------------------------
centroids <- st_centroid(adm3_boundaries_id)
centroids_coords <- st_coordinates(centroids)
centroids_df <- data.frame(
  node    = seq_len(nrow(adm3_boundaries_id)), 
  ADM2_ID = adm3_boundaries_id$ADM3_ID,        
  x       = centroids_coords[, 1],          
  y       = centroids_coords[, 2]       
)

# ------------------------------------------------------------------------------
# 2) Build the igraph object and extract edges
# ------------------------------------------------------------------------------
graph <- graph_from_adjacency_matrix(as.matrix(adj_matrix), mode = "undirected")
edges <- as_edgelist(graph)
edges_df <- data.frame(
  from = as.integer(edges[, 1]),
  to   = as.integer(edges[, 2])
) %>%
  left_join(centroids_df, by = c("from" = "node")) %>%
  rename(x = x, y = y) %>%
  left_join(centroids_df, by = c("to" = "node"), suffix = c("", "_end")) %>%
  rename(xend = x_end, yend = y_end)
ggplot() +
  geom_sf(
    data  = adm2_boundaries_id, 
    fill  = "white", 
    color = "black"
  ) +
  geom_segment(
    data = edges_df,
    aes(x = x, y = y, xend = xend, yend = yend),
    color = "red",
    alpha = 0.7
  ) +
  geom_point(
    data = centroids_df,
    aes(x = x, y = y),
    size = 2,
    alpha=0.6,
  ) +
  theme_void()

ggsave("nbstructure_admin3.png", 
       width = 13, height = 9, dpi = 400, bg = "white")

```



Repeat the above steps to estimate coverages at admin-3

```{r}
merged_data_sf <- st_join(data_ethiopia_1, drc1_sf, join = st_contains)merged_data_sf_id <- merged_data_sf %>%
  mutate(ADM3_ID = as.integer(as.factor(ADM3_EN)))
adm3_boundaries_id <- merged_data_sf_id %>%
  group_by(ADM3_ID) %>%
  slice(1) %>% 
  ungroup()
adm3_boundaries_id$ADM3_ID_iid <- adm3_boundaries_id$ADM3_ID

binom_df <- merged_data_sf_id %>% 
  group_by(cluster,ADM3_ID) %>% 
  summarise(y = sum(value, na.rm = T), 
            N = length(which(!(is.na(value)))),
            ADM3_EN = ADM3_EN %>% nth(1), 
            ADM3_ID = ADM3_ID %>% nth(1),
            urban = as.numeric(nth(strata, 1) == "urban"), 
            rural = as.numeric(nth(strata, 1) == "rural")) %>%
  as.data.frame()

binom_df$ADM3_ID_iid <- binom_df$ADM3_ID

nb_adm1 <- poly2nb(adm3_boundaries_id, row.names = adm3_boundaries_id$ADM3_ID)
adj_matrix <- nb2mat(nb_adm1, style = "B", zero.policy = TRUE)
adj_list_adm1 <- inla.read.graph(adj_matrix)
formula <- y ~ urban + rural - 1 +
  f(ADM3_ID, model = "bym2", graph = adj_list_adm1) + 
  f(ADM3_ID_iid, model = "iid")

binom_df$y[binom_df$N == 0] <- NA

result <- inla(formula = formula, 
               Ntrials = N,
               data = binom_df,
               family = "betabinomial",
               control.predictor = list(compute = TRUE),
               control.compute = list(return.marginals = TRUE,
                                      config = TRUE,
                                      return.marginals.predictor = TRUE))


#urban_prop is calculated from .tif file below

files <- c(
  "GHS_SMOD_E2030_GLOBE_R2023A_54009_1000_V2_0_R9_C23.tif",
  "GHS_SMOD_E2030_GLOBE_R2023A_54009_1000_V2_0_R9_C22.tif",
  "GHS_SMOD_E2030_GLOBE_R2023A_54009_1000_V2_0_R8_C23.tif",
  "GHS_SMOD_E2030_GLOBE_R2023A_54009_1000_V2_0_R8_C22.tif"
)
smod_list <- lapply(files, rast)

# Mosaic them into a single SpatRaster
# do.call(...) helps combine them
smod_mosaic <- do.call(mosaic, smod_list)
smod_mosaic

eth_adm3 <- adm3_boundaries_id
eth_adm3_moll <- st_transform(eth_adm3, crs(smod_mosaic))
eth_vect_moll <- vect(eth_adm3_moll)
smod_crop <- crop(smod_mosaic, eth_vect_moll)
smod_ethiopia <- mask(smod_crop, eth_vect_moll)
plot(smod_ethiopia)
reclass_mat <- matrix(c(
  10, 0,  # Water -> 0
  11, 0,  # Very low density rural -> 0
  12, 0,  # Low density rural -> 0
  13, 0,  # Rural cluster -> 0
  21, 1,  # Suburban/peri-urban -> 1
  22, 1,  # Semi-dense cluster -> 1
  23, 1,  # Dense urban cluster -> 1
  30, 1   # Urban centre -> 1
), ncol=2, byrow=TRUE)

#reclassify
urban_mask <- classify(smod_ethiopia, reclass_mat, include.lowest=FALSE, right=NA)

eth_adm3$urban_prop <- exact_extract(
  urban_mask, 
  eth_adm3,
  fun = function(values, coverage_fraction) {
    weighted.mean(values, coverage_fraction, na.rm = TRUE)
  }
)
urban_prop <- data.frame(
  urban_proportion <- eth_adm3$urban_prop,
  ADM3_ID <- eth_adm3$ADM3_ID,
  ADM3_EN <- eth_adm3$ADM3_EN
)


#urban variable is having lots of effect
# the estimation is being drawn to the mean value which is around 60%
# maybe the scarsity in the clsuter informaiton is leading to this

# ggplot(eth_adm3) +
#   geom_sf(aes(fill = urban_prop), color = "white", size=0.2) +
#   scale_fill_viridis_c(option="C", name="Urban Fraction") +
#   labs(title="Urban Proportion by ADM2 (GHSL SMOD)") +
#   theme_minimal()

# urban_prop <- binom_df %>%
#   group_by(ADM2_EN, ADM2_ID) %>%
#   summarize(
#     total_observations = n(),
#     urban_count = sum(urban == 1, na.rm = TRUE),
#     urban_proportion = mean(urban == 1, na.rm = TRUE)
#   ) %>%
#   ungroup()
urban_prop$urban_proportion[is.nan(urban_prop$urban_proportion)] <- 0
nsamp <- 1000
samp <- inla.posterior.sample(n = nsamp, result = result)
region_idx <- which(rownames(samp[[1]]$latent) %>% 
                      str_detect("ADM3"))[1:length(unique(binom_df$ADM3_ID))]

urban_idx <- which(rownames(samp[[1]]$latent) %>% str_detect("urban"))
rural_idx <- which(rownames(samp[[1]]$latent) %>% str_detect("rural"))

region_mat <- matrix(0, nrow = length(region_idx), ncol = nsamp)
urban_mat <- matrix(0, nrow = 1, ncol = nsamp)
rural_mat <- matrix(0, nrow = 1, ncol = nsamp)
for (i in 1:nsamp) {
  region_mat[,i] <- samp[[i]]$latent[region_idx]
  urban_mat[,i] <- samp[[i]]$latent[urban_idx]
  rural_mat[,i] <- samp[[i]]$latent[rural_idx]
}
samp_mat_urban <- region_mat + urban_mat[rep(1, nrow(region_mat)),]
samp_mat_rural <- region_mat + rural_mat[rep(1, nrow(region_mat)),]
expit_samp_mat_urban <- expit(samp_mat_urban)
expit_samp_mat_rural <- expit(samp_mat_rural)
expit_samp_mat <- (urban_prop$urban_proportion * expit_samp_mat_urban + (1-urban_prop$urban_proportion) * expit_samp_mat_rural)

samp_betabinom <- expit_samp_mat


bino <- data.frame(
  ADM3_ID = sort(unique(adm3_boundaries_id$ADM3_ID)),
  ADM3_EN = sort(unique(adm3_boundaries_id$ADM3_EN)),
  med_beta_bino = apply(samp_betabinom, 1, median),
  lower_beta_bino = apply(samp_betabinom, 1, quantile, 0.025),
  upper_beta_bino = apply(samp_betabinom, 1, quantile, 0.975),
  sd_beta_bino = apply(samp_betabinom, 1, sd)
)
plot_plot_adm3 <- left_join(adm3_boundaries_id, bino, by="ADM3_ID")

plot_plot_adm3$med_beta_bino <- plot_plot_adm3$med_beta_bino * 100
plot_plot_adm3$sd_beta_bino <- plot_plot_adm3$sd_beta_bino * 100

st_write(plot_plot_adm3, "result/unit_level_estimate_admin3_urban_new.shp")
# Create polished visualization
ggplot() +
  geom_sf(data = subset(plot_plot, !is.na(med_beta_bino) & sd_beta_bino != 0), 
          aes(fill = med_beta_bino), color = "white", lwd = 0.05) +
  scale_fill_viridis(
    name = "Measles Coverage (%)",
    option = "mako",  
    direction = -1,
    begin = 0.15,
    end = 0.85,
    na.value = "transparent",
    limits = c(0, 100),
    guide = guide_colorbar(
      direction = "horizontal",
      barheight = unit(2, "mm"),
      barwidth = unit(80, "mm"),
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "gray30",
      ticks.colour = "gray30")
  ) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = subset(plot_plot, sd_beta_bino == 0),
          fill = "#E63946", color = "white", lwd = 0.05) +  
  geom_sf(data = subset(plot_plot, is.na(med_beta_bino)),
          fill = "#495057", color = "white", lwd = 0.05) +  
   geom_sf(
  data = adm1_boundaries,
  fill = NA,
  color = "black",
  alpha = 0.7,
  linewidth = 1.1
) +
#NEW layer right after existing admin1 layer

  # Scale bar with refined style
  # annotation_scale(
  #   location = "bl",
  #   style = "ticks",
  #   line_width = 0.4,
  #   text_cex = 0.9,
  #   height = unit(0.2, "cm"),
  #   pad_x = unit(0.2, "cm"),
  #   pad_y = unit(0.2, "cm")
  # ) +
  theme_void() +
  theme(
    # plot.background = element_rect(fill = "#F8F9FA", color = NA),
    # panel.background = element_rect(fill = "#F8F9FA", color = NA),
    legend.position = "bottom",
    legend.title = element_text(
      size = 20,
      face = "bold",
      margin = margin(b = 3)),
    legend.text = element_text(size = 17),
  
    plot.title = element_text(
      face = "bold",
      size = 20,
      hjust = 0.5,
      color = "#2C3E50",
      margin = margin(b = 8)),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      color = "#7F8C8D",
      margin = margin(b = 15)),
    plot.caption = element_text(
      size = 9,
      color = "#95A5A6",
      hjust = 0.98,
      margin = margin(t = 10))
  ) +
  labs(
    title = "",
    subtitle = "",
    caption = ""
  ) 


#here we have used urban prop form GSHL
ggsave("measles_coverage_final_admin3_cluster_level_new.png", 
       width = 16, height = 12, dpi = 400, bg = "white")
```


Admin 3 no urban and rural considerations



```{r}

merged_data_sf_id <- merged_data_sf %>%
  mutate(ADM3_ID = as.integer(as.factor(ADM3_EN)))

adm3_boundaries_id <- merged_data_sf_id %>%
  group_by(ADM3_ID) %>%
  slice(1) %>%
  ungroup()

adm3_boundaries_id$ADM3_ID_iid <- adm3_boundaries_id$ADM3_ID

binom_df <- merged_data_sf_id %>%
  group_by(cluster, ADM3_ID) %>%
  summarise(
    y       = sum(value, na.rm = TRUE),
    N       = sum(!is.na(value)),
    ADM3_EN = nth(ADM3_EN, 1),
    ADM3_ID = nth(ADM3_ID, 1),
    .groups = "drop"
  ) %>%
  as.data.frame()

binom_df$ADM3_ID_iid <- binom_df$ADM3_ID
binom_df$y[binom_df$N == 0] <- NA

nb_adm3 <- poly2nb(adm3_boundaries_id, row.names = adm3_boundaries_id$ADM3_ID)
adj_matrix <- nb2mat(nb_adm3, style = "B", zero.policy = TRUE)
adj_list_adm3 <- inla.read.graph(adj_matrix)

formula <- y ~ -1 + 
  f(ADM3_ID, model = "bym2", graph = adj_list_adm3) +
  f(ADM3_ID_iid, model = "iid")

result <- inla(
  formula         = formula,
  family          = "betabinomial",
  Ntrials         = binom_df$N,
  data            = binom_df,
  control.predictor = list(compute = TRUE),
  control.compute   = list(
    return.marginals           = TRUE,
    config                     = TRUE,
    return.marginals.predictor = TRUE
  )
)

nsamp <- 1000
samp  <- inla.posterior.sample(n=nsamp, result=result)

region_idx <- which(rownames(samp[[1]]$latent) %>% 
                      str_detect("ADM3"))[1:length(unique(binom_df$ADM3_ID))]

region_names <- rownames(samp[[1]]$latent)[region_idx]

region_mat <- matrix(0, nrow = length(region_idx), ncol = nsamp)

for (i in seq_len(nsamp)) {
  region_mat[, i] <- samp[[i]]$latent[region_idx]
}

combined_mat <- region_mat

n_region <- length(region_idx)
samp_mat <- matrix(0, nrow=n_region, ncol=nsamp)

for (r in seq_len(n_region)) {
  for (i in seq_len(nsamp)) {
    samp_mat[r, i] <- combined_mat[r, i]
  }
}

expit_samp_mat <- 1 / (1 + exp(-samp_mat))

med_vals   <- apply(expit_samp_mat, 1, median)
lower_vals <- apply(expit_samp_mat, 1, quantile, 0.025)
upper_vals <- apply(expit_samp_mat, 1, quantile, 0.975)
sd_vals    <- apply(expit_samp_mat, 1, sd)

bino <- data.frame(
  ADM3_ID        = adm3_boundaries_id$ADM3_ID,
  med_beta_bino  = med_vals,
  lower_beta_bino= lower_vals,
  upper_beta_bino= upper_vals,
  sd_beta_bino   = sd_vals
)

plot_plot_adm3_no <- left_join(adm3_boundaries_id, bino, by="ADM3_ID")

plot_plot_adm3_no$med_beta_bino <- 100 * plot_plot_adm3_no$med_beta_bino
plot_plot_adm3_no$sd_beta_bino  <- 100 * plot_plot_adm3_no$sd_beta_bino

st_write(plot_plot_adm3_no, "result/unit_level_estimate_admin3_no_urban.shp")

ggplot() +
  geom_sf(data = subset(plot_plot, !is.na(med_beta_bino) & sd_beta_bino != 0), 
          aes(fill = med_beta_bino), color = "white", lwd = 0.05) +
  scale_fill_viridis(
    name = "Measles Coverage (%)",
    option = "mako",
    direction = -1,
    begin = 0.15,
    end = 0.85,
    na.value = "transparent",
    limits = c(0, 100),
    guide = guide_colorbar(
      direction = "horizontal",
      barheight = unit(2, "mm"),
      barwidth = unit(80, "mm"),
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "gray30",
      ticks.colour = "gray30"
    )
  ) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = subset(plot_plot, sd_beta_bino == 0),
          fill = "#E63946", color = "white", lwd = 0.05) +
  geom_sf(data = subset(plot_plot, is.na(med_beta_bino)),
          fill = "#495057", color = "white", lwd = 0.05) +
  geom_sf(
    data = adm1_boundaries,
    fill = NA,
    color = "black",
    alpha = 0.7,
    linewidth = 1.1
  ) +
  theme_void() +
  theme(
    legend.position    = "bottom",
    legend.title       = element_text(size = 20, face = "bold", margin = margin(b = 3)),
    legend.text        = element_text(size = 17),
    plot.title         = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle      = element_text(size = 11, color = "grey30", hjust = 0.5, margin = margin(b = 15)),
    plot.caption       = element_text(size = 9, color = "#95A5A6", hjust = 0.98, margin = margin(t = 10))
  )

ggsave("plot_new.png", 
       width = 16, height = 12, dpi = 400, bg = "white")

# Second administrative-level model without urban/rural
merged_data_sf_id <- merged_data_sf %>%
  mutate(ADM2_ID = as.integer(as.factor(ADM2_EN)))

adm2_boundaries_id <- merged_data_sf_id %>%
  group_by(ADM2_EN, ADM2_ID) %>%
  summarize(geometry = st_union(geometry)) %>%
  st_as_sf()

adm2_boundaries_id$ADM2_ID_iid <- adm2_boundaries_id$ADM2_ID

binom_df <- merged_data_sf_id %>%
  group_by(cluster, ADM2_ID) %>%
  summarise(
    y       = sum(value, na.rm=TRUE),
    N       = sum(!is.na(value)),
    ADM2_EN = nth(ADM2_EN, 1),
    ADM2_ID = nth(ADM2_ID, 1),
    .groups = "drop"
  ) %>%
  as.data.frame()

binom_df$ADM2_ID_iid <- binom_df$ADM2_ID
binom_df$y[binom_df$N == 0] <- NA

nb_adm1 <- poly2nb(binom_sf, row.names = binom_sf$ADM2_ID)
adj_matrix <- nb2mat(nb_adm1, style="B", zero.policy=TRUE)
adj_list_adm1 <- inla.read.graph(adj_matrix)

formula <- y ~ -1 + 
  f(ADM2_ID, model="bym2", graph=adj_list_adm1) + 
  f(ADM2_ID_iid, model="iid")

result <- inla(
  formula         = formula,
  Ntrials         = binom_df$N,
  data            = binom_df,
  family          = "betabinomial",
  control.predictor = list(compute=TRUE),
  control.compute   = list(
    return.marginals           = TRUE,
    config                     = TRUE,
    return.marginals.predictor = TRUE
  )
)

nsamp <- 1000
samp  <- inla.posterior.sample(n=nsamp, result=result)

region_idx <- which(rownames(samp[[1]]$latent) %>% 
                      str_detect("ADM2"))[1:length(unique(binom_df$ADM2_ID))]

region_mat <- matrix(0, nrow=length(region_idx), ncol=nsamp)

for(i in seq_len(nsamp)) {
  region_mat[, i] <- samp[[i]]$latent[region_idx]
}

linpred_mat <- region_mat
expit_samp_mat <- 1 / (1 + exp(-linpred_mat))

bino <- data.frame(
  ADM2_ID         = sort(unique(adm2_boundaries_id$ADM2_ID)),
  ADM2_EN         = sort(unique(adm2_boundaries_id$ADM2_EN)),
  med_beta_bino   = apply(expit_samp_mat, 1, median),
  lower_beta_bino = apply(expit_samp_mat, 1, quantile, 0.025),
  upper_beta_bino = apply(expit_samp_mat, 1, quantile, 0.975),
  sd_beta_bino    = apply(expit_samp_mat, 1, sd)
)

plot_plot_no <- left_join(adm2_boundaries_id, bino, by="ADM2_ID")

plot_plot_no$med_beta_bino <- plot_plot_no$med_beta_bino * 100
plot_plot_no$sd_beta_bino  <- plot_plot_no$sd_beta_bino  * 100

st_write(plot_plot_no, "result/unit_level_estimate_admin2_no_urban.shp")

ggplot() +
  geom_sf(data = subset(plot_plot, !is.na(med_beta_bino) & sd_beta_bino != 0), 
          aes(fill = med_beta_bino), color = "white", lwd = 0.05) +
  scale_fill_viridis(
    name = "Measles Coverage (%)",
    option = "mako", 
    direction = -1,
    begin = 0.15,
    end   = 0.85,
    na.value = "transparent",
    limits = c(0, 100),
    guide = guide_colorbar(
      direction="horizontal",
      barheight=unit(2,"mm"),
      barwidth=unit(80,"mm"),
      title.position="top",
      title.hjust=0.5,
      frame.colour="gray30",
      ticks.colour="gray30"
    )
  ) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = subset(plot_plot, sd_beta_bino==0),
          fill="#E63946", color="white", lwd=0.05) +
  geom_sf(data = subset(plot_plot, is.na(med_beta_bino)),
          fill="#495057", color="white", lwd=0.05) +
  geom_sf(
    data = adm1_boundaries,
    fill = NA,
    color = "black",
    alpha = 0.7,
    linewidth = 1.1
  ) +
  theme_void() +
  theme(
    legend.position="bottom",
    legend.title=element_text(size=20, face="bold", margin=margin(b=3)),
    legend.text=element_text(size=17),
    plot.title=element_text(face="bold", size=20, hjust=0.5, color="#2C3E50", margin=margin(b=8)),
    plot.subtitle=element_text(size=12, hjust=0.5, color="#7F8C8D", margin=margin(b=15)),
    plot.caption=element_text(size=9, color="#95A5A6", hjust=0.98, margin=margin(t=10))
  )

ggsave("measles_coverage_final_admin2_cluster_level_NO_URBAN_RURAL.png", 
       width=16, height=12, dpi=400, bg="white")


```


Admin2 accuracy assessment will follow:

first, read all

```{r}
all_data <- st_read("shapefile_containing_individual_and _cluster_level_data.shp")

est_2_direct <- 
  st_read("result/direct_estimate_admin2.shp")

est_2_direct$se <- est_2_direct$se * 100
est_2_direct$value <- est_2_direct$value * 100

est_2_unit_level_urban <- 
  st_read("result/unit_level_estimate_admin3_urban.shp")


est_2_unit_level_no_urban <- 
  st_read("result/unit_level_estimate_admin2_no_urban.shp")
direct <- est_2_direct %>% 
  st_drop_geometry() %>% 
  select(se) %>% 
  rename(Direct_SE = se)

urban <- est_2_unit_level_urban %>% 
  st_drop_geometry() %>% 
  select(sd_bt_b) %>% 
  rename(Urban_SE = sd_bt_b)

no_urban <-  est_2_unit_level_no_urban %>%  # Note: Verify path for actual no_urban data
  st_drop_geometry() %>% 
  select(sd_bt_b) %>% 
  rename(NoUrban_SE = sd_bt_b)
combined <- bind_cols(direct, urban, no_urban) %>% 
  pivot_longer(cols = c(Urban_SE, NoUrban_SE),
               names_to = "Method", 
               values_to = "Unit_SE")
ggplot(combined, aes(x = Direct_SE, y = Unit_SE, color = Method, shape = Method)) +
  geom_point(alpha = 0.9, size = 3, shape=21) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey40") +
  scale_color_manual(values = c("Urban_SE" = "#E69F00", "NoUrban_SE" = "#56B4E9"),
                     labels = c("Urban Unit-Level", "Non-Urban Unit-Level")) +
  scale_shape_manual(values = c("Urban_SE" = 17, "NoUrban_SE" = 19),
                    labels = c("Urban Unit-Level", "Non-Urban Unit-Level")) +
  labs(title = "",
       subtitle = "",
       x = "Direct Estimate Standard Error",
       y = "Unit-Level Model Standard Deviation",
       caption = "") +
  theme_pubclean()  +
  theme(
    plot.title = element_text( hjust = 0.5, size = 18),
    axis.title = element_text(size = 14),
    panel.grid = element_line(color = "#EAEDED"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white", color = NA),
    axis.line = element_line(color = "#2C3E50"),
    plot.caption = element_text(hjust = 0, color = "#7F8C8D")
  ) +
  coord_cartesian(ylim = c(0, 22)) +
  scale_x_continuous(expand = expansion(mult = 0.02)) +
  scale_y_continuous(expand = expansion(mult = 0.02))

```


```{r}
urban <- est_2_unit_level_urban %>%
  st_drop_geometry() %>%
  select(urban_se = sd_bt_b)

no_urban <- est_2_unit_level_no_urban %>%
  st_drop_geometry() %>%
  select(nourban_se = sd_bt_b)
combined <- bind_cols(urban, no_urban)
ggplot(combined, aes(x = urban_se, y = nourban_se)) +
  geom_hex(bins = 40, show.legend = FALSE) +
  geom_point(alpha = 0.4, color = "#2E86C1", size = 3) +
  geom_smooth(method = "lm", color = "#E74C3C", se = FALSE, 
              linewidth = 1.2, linetype = "dotdash") +
  geom_abline(intercept = 0, slope = 1, color = "#27AE60", 
              linewidth = 1, linetype = "longdash") +
  stat_cor(method = "pearson", label.x.npc = 0.05, label.y.npc = 0.95,
           aes(label = ..r.label..), color = "#34495E") +
  scale_fill_gradient(low = "#D6EAF8", high = "#2E86C1") +
  annotate("text", x = 10, y = 15, label = "Reference",
           color = "#27AE60", size = 4.5, hjust = 1, fontface = "bold") +
  annotate("text", x = 10, y = 17, label = "Linear trend",
           color = "#E74C3C", size = 4.5, hjust = 1, fontface = "bold") +
  coord_equal(xlim = c(0, 23), ylim = c(0, 23)) +
  labs(title = "",
       x = "Urban Standard Deviation",
       y = "Non-Urban Standard Deviation",
       caption = "") +
  theme_pubclean() +
  theme(
    plot.title = element_text( hjust = 0.5, size = 18),
    axis.title = element_text(size = 14),
    panel.grid = element_line(color = "#EAEDED"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "white", color = NA),
    axis.line = element_line(color = "#2C3E50"),
    plot.caption = element_text(hjust = 0, color = "#7F8C8D")
  )
```

```{r}
ggsave("direct_cluster_level_NO_URBAN_RURAL_SE.png", 
       width=8, height=5, dpi=400, bg="white")
```


```{r}
cluster_agg <- merged_data_sf %>%
  st_drop_geometry() %>%
  group_by(cluster) %>%
  summarise(n_sample = sum(!is.na(value)))
cluster_points <- drc1_sf %>%
  group_by(cluster) %>%
  slice(1) %>%
  ungroup()
cluster_counts <- left_join(cluster_points, cluster_agg, by = "cluster")
adm2_boundaries <- data_ethiopia_1 %>%
  group_by(ADM2_EN) %>%
  summarize(geometry = st_union(geometry)) %>%
  st_as_sf()
adm3_sample_data <- cluster_counts %>%
  st_join(
    data_ethiopia_1 %>% select(ADM3_EN),  # Spatial join clusters to ADM3 polygons
    join = st_within
  ) %>%
  st_drop_geometry() %>%
  group_by(ADM3_EN) %>%
  summarise(total_samples = sum(n_sample, na.rm = TRUE)) %>%
  right_join(data_ethiopia_1, by = "ADM3_EN") %>%  # Join back to original polygons
  st_as_sf()



adm2_sample_data <- cluster_counts %>%
  st_join(
    adm2_boundaries %>% select(ADM2_EN),  
    join = st_within
  ) %>%
  st_drop_geometry() %>%
  group_by(ADM2_EN) %>%
  summarise(total_samples = sum(n_sample, na.rm = TRUE)) %>%
  right_join(adm2_boundaries, by = "ADM2_EN") %>% 
  st_as_sf()

adm2_sample_data <- adm2_sample_data %>% st_drop_geometry()
direct <- est_2_direct %>% 
  st_drop_geometry() %>% 
  select(se, ADM2_EN) %>% 
  rename(Direct_SE = se)

urban <- est_2_unit_level_urban %>% 
  st_drop_geometry() %>% 
  select(sd_bt_b, ADM2_EN_x) %>% 
  rename(Urban_SE = sd_bt_b,
         ADM2_EN = ADM2_EN_x)

no_urban <-  est_2_unit_level_no_urban %>%  # Note: Verify path for actual no_urban data
  st_drop_geometry() %>% 
  select(sd_bt_b, ADM2_EN_x) %>% 
  rename(No_Urban_SE = sd_bt_b,
         ADM2_EN = ADM2_EN_x)

merged_data_plot <- adm2_sample_data %>%
  full_join(direct, by = "ADM2_EN") %>%
  full_join(urban, by = "ADM2_EN") %>%
  full_join(no_urban, by = "ADM2_EN")

merged_data_plot$abs_de_urb <- abs(merged_data_plot$Direct_SE - merged_data_plot$Urban_SE)
merged_data_plot$abs_de_no_urb <- abs(merged_data_plot$Direct_SE - merged_data_plot$No_Urban_SE)



```


```{r}


merged_data_plot <- merged_data_plot %>% filter(
  total_samples <= 23
)
ggplot(merged_data_plot, aes(x = total_samples)) +
  geom_point(aes(y = abs_de_urb, color = "Urban Model", size=abs_de_urb), 
             alpha = 0.8,  shape = 21) +
  geom_point(aes(y = abs_de_no_urb, color = "Non-Urban Model", size=abs_de_no_urb), 
             alpha = 0.8, size = 2.5, shape = 21) +
  geom_line(stat = "smooth",aes(y = abs_de_urb, color = "Urban Model"), 
              method = "gam", formula = y ~ s(x, bs = "cs"),
              se = FALSE, linewidth = 1.5, alpha=0.6) +
  geom_line(stat="smooth", aes(y = abs_de_no_urb, color = "Non-Urban Model"), 
              method = "gam", formula = y ~ s(x, bs = "cs"),
              se = FALSE, linewidth = 1.5, linetype = "dashed", alpha=0.6) +
  geom_hline(yintercept = 0, color = "grey50", linetype = "dotted") +
  geom_text(x = max(merged_data_plot$total_samples, na.rm = TRUE)*0.9, y = 9,
            label = "Urban Model", color = "#D35400", 
            hjust = 1, fontface = "bold", size = 4.5) +
  geom_text(x = max(merged_data_plot$total_samples, na.rm = TRUE)*0.9, y = 8,
            label = "Non-Urban Model", color = "#2980B9", 
            hjust = 1, fontface = "bold", size = 4.5) +
  scale_color_manual(values = c("Urban Model" = "#D35400", 
                               "Non-Urban Model" = "#2980B9")) +
  scale_x_continuous(
    name = "Sample Size",
    breaks = scales::breaks_extended(8),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  scale_y_continuous(
    name = "Absolute SE Difference (|Direct - Model|)",
    limits = c(0, 23),
    breaks = seq(0, 23, 2),
    expand = expansion(mult = c(0, 0.02))
  ) +
  labs(
    title = "",
    subtitle = ""
  ) +
  theme_pubclean()+
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "grey30",
                                margin = margin(b = 15)),
    axis.title = element_text(face = "bold", size = 12),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text = element_text(size = 10, color = "grey40"),
    axis.ticks = element_line(color = "grey60", linewidth = 0.6),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "none"
  ) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf,
           color = "grey40", linewidth = 0.8) +
  annotate("segment", y = -Inf, yend = Inf, x = -Inf, xend = -Inf,
           color = "grey40", linewidth = 0.8)
```

```{r}
ggsave("shrinkage.png", 
       width=8, height=5, dpi=400, bg="white")
```


Leave one area out cross validation.. as follows still working at admin2


```{r}
expit <- function(x) {
  1 / (1 + exp(-x))
}

# -------------------------------
# 
#   - binom_df: unit-level data with columns ADM2_ID, ADM2_EN, y, N, urban, rural
#   - formula: the INLA formula
#   - urban_prop: a data frame with urban proportion per area, with columns:
#         urban_proportion, ADM2_ID, ADM2_EN
#   - est_2_direct: direct estimates with columns ADM2_EN, value, se
# -------------------------------
urban_prop$urban_proportion[is.nan(urban_prop$urban_proportion)] <- 0
nsamp <- 1000
unique_ids <- unique(binom_df$ADM2_ID)

# -------------------------------
# 1) Leave-One-Area Cross Validation using Posterior Sampling
# -------------------------------
loo_results <- data.frame()
for (area in unique_ids) {
  cat("Processing LOOCV for area:", area, "\n")
  loo_data <- na.omit(binom_df)
  loo_data$y[loo_data$ADM2_ID == area] <- NA
  result_loo <- inla(
    formula = formula,
    Ntrials = loo_data$N,
    data = loo_data,
    family = "betabinomial",
    control.predictor = list(compute = TRUE),
    control.compute = list(return.marginals = TRUE,
                           config = TRUE,
                           return.marginals.predictor = TRUE)
  )
  samp <- inla.posterior.sample(n = nsamp, result = result_loo)
  region_idx <- which(str_detect(rownames(samp[[1]]$latent), "ADM2"))[1:length(unique_ids)]
  urban_idx <- which(str_detect(rownames(samp[[1]]$latent), "urban"))
  rural_idx <- which(str_detect(rownames(samp[[1]]$latent), "rural"))
  region_mat <- matrix(0, nrow = length(region_idx), ncol = nsamp)
  urban_mat  <- matrix(0, nrow = 1, ncol = nsamp)
  rural_mat  <- matrix(0, nrow = 1, ncol = nsamp)
  for(i in 1:nsamp) {
    region_mat[, i] <- samp[[i]]$latent[region_idx]
    urban_mat[, i]  <- samp[[i]]$latent[urban_idx]
    rural_mat[, i]  <- samp[[i]]$latent[rural_idx]
  }
  samp_mat_urban <- region_mat + matrix(rep(urban_mat, each = nrow(region_mat)), nrow = nrow(region_mat))
  samp_mat_rural <- region_mat + matrix(rep(rural_mat, each = nrow(region_mat)), nrow = nrow(region_mat))
  expit_samp_mat_urban <- expit(samp_mat_urban)
  expit_samp_mat_rural <- expit(samp_mat_rural)
  u_prop <- urban_prop$urban_proportion[urban_prop$ADM2_ID == area]
  if(length(u_prop) > 1) { u_prop <- u_prop[1] }
  expit_samp_mat <- (u_prop * expit_samp_mat_urban + (1 - u_prop) * expit_samp_mat_rural)

  pos <- which(unique_ids == area)
  pred_samples <- expit_samp_mat[pos, ]
  med   <- median(pred_samples)
  lower <- as.numeric(quantile(pred_samples, 0.025))
  upper <- as.numeric(quantile(pred_samples, 0.975))
  sd_val <- sd(pred_samples)
  adm2_en <- loo_data$ADM2_EN[match(area, loo_data$ADM2_ID)]
  loo_results <- rbind(loo_results, data.frame(
    ADM2_ID    = area,
    ADM2_EN    = adm2_en,
    pred_mean  = med,
    pred_lower = lower,
    pred_upper = upper,
    sd_beta_bino = sd_val
  ))
}

# -------------------------------
# 2) Prepare Direct Estimates for Comparison
# -------------------------------
est_2_direct_good <- design_based_estimates_admin2 %>%
  filter(!is.na(value), se > 0.005)

# -------------------------------
# 3) Merge LOOCV Predictions with Direct Estimates
# -------------------------------
cv_df <- left_join(est_2_direct_good, loo_results, by = "ADM2_EN")
cv_df$pred_mean <- cv_df$pred_mean * 100
# -------------------------------
# 4) Plot Direct vs. LOOCV Predictions
# -------------------------------

p <- ggplot(cv_df, aes(x = value*100, y = pred_mean*100)) +

    geom_point(aes(size = se), color="#D35400", shape=21, alpha = 0.85, show.legend = TRUE) +
  
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "#3a86ff", linewidth = 1) +
  scale_color_viridis(
    option = "plasma", direction = -1, name = "Prediction SE"
  ) +
  # scale_size_continuous(name = "Prediction SE", range = c(2, 8)) +
  labs(
    x = "Direct Estimate",
    y = "LOOCV Model Prediction",
    title = "",
    subtitle = "",
    caption = ""
  ) +
 stat_cor(method = "pearson", label.x.npc = 0.05, label.y.npc = 0.95,
           aes(label = ..r.label..), color = "#34495E") +
  theme_pubclean()+
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "grey30",
                                margin = margin(b = 15)),
    axis.title = element_text(face = "bold", size = 40),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text = element_text(size = 34, color = "grey40"),
    axis.ticks = element_line(color = "grey60", linewidth = 0.6),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "none"
  ) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf,
           color = "grey40", linewidth = 0.8) +
  annotate("segment", y = -Inf, yend = Inf, x = -Inf, xend = -Inf,
           color = "grey40", linewidth = 0.8)
print(p)


rmse <- sqrt(mean((cv_df$value - cv_df$pred_mean)^2, na.rm = TRUE))
print(rmse)
rmse
write.csv(cv_df, "result/loocv_admin2.csv")
```

rmse for urban was 20.6.. and for when not using urban 27.9

```{r}
ggsave("cross_validated.png", 
       width=6, height=4, dpi=400, bg="white")
```


```{r}
direct <- est_2_direct %>% 
  st_drop_geometry() %>% 
  select(value, se, ADM2_EN) %>% 
  rename(Direct_value = value)

urban <- est_2_unit_level_urban %>% 
  st_drop_geometry() %>% 
  select(md_bt_b, ADM2_EN_x) %>% 
  rename(Urban_MD = md_bt_b)

bb <- left_join(direct, urban, by=c("ADM2_EN" = "ADM2_EN_x"))


p <- ggplot(ccc, aes(x = value.x*100, y = med_beta_bino)) +

    geom_point(aes(size=sd_beta_bino),color="#D35400", shape=21, alpha = 0.85, show.legend = TRUE) +
  
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "#3a86ff", linewidth = 1) +
  scale_color_viridis(
    option = "plasma", direction = -1, name = "Prediction SE"
  ) +
  # scale_size_continuous(name = "Prediction SE", range = c(2, 8)) +
  labs(
    x = "Direct Estimate",
    y = "Model Based (Using Urban)",
    title = "",
    subtitle = "",
    caption = ""
  ) +
  stat_cor(method = "pearson", label.x.npc = 0.05, label.y.npc = 0.95,
           aes(label = ..r.label..), color = "#34495E", size=14) +
  theme_pubclean()+
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "grey30",
                                margin = margin(b = 15)),
    axis.title = element_text(face = "bold", size = 15),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text = element_text(size = 15, color = "grey40"),
    axis.ticks = element_line(color = "grey60", linewidth = 0.6),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "none"
  ) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf,
           color = "grey40", linewidth = 0.8) +
  annotate("segment", y = -Inf, yend = Inf, x = -Inf, xend = -Inf,
           color = "grey40", linewidth = 0.8)
print(p)
```


```{r}
ggsave("no_cross_validated_coapring_dire_model_based_mean_admin3.png", 
       width=6, height=4, dpi=400, bg="white")
```


```{r}
direct <- est_2_direct %>% 
  st_drop_geometry() %>% 
  select(value, se, ADM3_EN) %>% 
  rename(Direct_value = value)

urban <- est_2_unit_level_urban %>% 
  st_drop_geometry() %>% 
  select(md_bt_b, ADM3_EN_x) %>% 
  rename(Urban_MD = md_bt_b)

bb <- left_join(direct, urban, by=c("ADM3_EN" = "ADM3_EN_x"))


p <- ggplot(bb, aes(x = Direct_value, y = Urban_MD)) +

    geom_point(aes(size=se),color="#D35400", shape=21, alpha = 0.85, show.legend = TRUE) +
  
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "#3a86ff", linewidth = 1) +
  scale_color_viridis(
    option = "plasma", direction = -1, name = "Prediction SE"
  ) +
  # scale_size_continuous(name = "Prediction SE", range = c(2, 8)) +
  labs(
    x = "Direct Estimate",
    y = "Model Based (Using Urban)",
    title = "",
    subtitle = "",
    caption = ""
  ) +
  stat_cor(method = "pearson", label.x.npc = 0.05, label.y.npc = 0.95,
           aes(label = ..r.label..), color = "#34495E", size=4) +
  theme_pubclean()+
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "grey30",
                                margin = margin(b = 15)),
    axis.title = element_text(face = "bold", size = 17),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text = element_text(size = 14, color = "grey40"),
    axis.ticks = element_line(color = "grey60", linewidth = 0.6),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "none"
  ) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf,
           color = "grey40", linewidth = 0.8) +
  annotate("segment", y = -Inf, yend = Inf, x = -Inf, xend = -Inf,
           color = "grey40", linewidth = 0.8)
print(p)
```

```{r}
ggsave("no_cross_validated_coapring_dire_model_based_mean_admin3.png", 
       width=6, height=4, dpi=400, bg="white")
```


For writing description in the poster!

```{r}
urba_2 <- st_read("result/unit_level_estimate_admin3_urban.shp")

no_urba_2 <- st_read("result/unit_level_estimate_admin3_no_urban.shp")
names(no_urba_2)

```


# more analysis

admin2 (these variables came from the above codes)

```{r}
adm2_sample_data_no_shape <- st_drop_geometry(adm2_sample_data)

merge1 <- left_join(adm2_sample_data_no_shape, design_based_estimates_admin2, by=c("ADM2_EN"))

merge2 <- left_join(merge1, adm2_summary, by=c("ADM2_EN"))

plot2 <- ggplot(na.omit(merge2), aes(x = total_samples, y = se, color = homogeneous)) +
  geom_point(size = 3) +
  geom_text_repel(
    aes(label = total_samples),
    size = 3,
    show.legend = FALSE,
    max.overlaps = 20
  ) +
  scale_color_brewer(
    palette = "Dark2",
    labels = c("FALSE" = "Heterogeneous", "TRUE" = "Homogeneous")
  ) +
  labs(
    x     = expression(paste("Total Samples per ", 2^{nd}, " administrative Unit")),
    y     = "Design-based Standard Error",
    color = "Response Homogeneity"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )





```

```{r}
design_based_estimates_admin3 <- design_based_estimates
adm3_sample_data_no_shape <- st_drop_geometry(adm3_sample_data)

merge1 <- left_join(adm3_sample_data_no_shape, design_based_estimates_admin3, by=c("ADM3_EN"))

merge3 <- left_join(merge1, adm3_summary, by=c("ADM3_EN"))
plot1 <- ggplot(merge3, aes(x = total_samples, y = se, color = homogeneous)) +
  geom_point(size = 3) +
  geom_text_repel(
    aes(label = total_samples),
    size = 3,
    show.legend = FALSE,
    max.overlaps = 20
  ) +
  scale_color_brewer(
    palette = "Dark2",
    labels = c("FALSE" = "Heterogeneous", "TRUE" = "Homogeneous")
  ) +
  labs(
    x     = expression(paste("Total Samples per ", 3^{rd}, " administrative Unit")),
    y     = "Design-based Standard Error",
    color = "Response Homogeneity"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

```


```{r}
final_plot <- plot1 + plot2 +
  plot_layout(ncol = 2) 
print(final_plot)

ggsave("SSvsSE.png", 
       width = 13, height = 7.4, dpi = 400, bg = "white")
```




```{r}
merge2 <- merge2 %>% filter(
  !is.na(homogeneous)
)
merge3 <- merge3 %>% filter(
  !is.na(homogeneous)
)
plot1 <- ggplot(merge2, aes(x = total_samples, y = se)) +
  geom_hex(bins = 30) +
  scale_fill_distiller(
    palette = "Spectral",
    direction = 1,
    name = "Count"
  ) +
  facet_wrap(~ homogeneous,
             labeller = as_labeller(c(
               "TRUE"  = "Homogeneous",
               "FALSE" = "Heterogeneous"
             ))) +
  labs(
    x     = expression(paste("Total Samples per ", 2^{nd}, " administrative Unit")),
    y     = "Design-based Standard Error",
    title = ""
  ) +
  theme_bw(base_size = 14) +
  theme(
    
    text = element_text(family = "Arial Narrow", color = "#2a2a2a"),
        plot.title = element_markdown(hjust = 0.5, lineheight = 1.2, 
                                      margin = margin(t = 10, b = 10)),
        plot.subtitle = element_markdown(hjust = 0.5, size = 16,
                                         color = "gray30", margin = margin(b = 20)),
        axis.title = element_text(face = "bold", size = 16),
        legend.position = "bottom",
        legend.text = element_text(size = 13),
        axis.text.x = element_text(angle = 90, hjust = 1, color = "gray40"),
        axis.text.y = element_text(color = "gray40"),
        # panel.grid.major.x = element_blank(),
        # panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        strip.background = element_rect(fill = "#F7F7F7", color = NA),
        strip.text = element_text(face="bold", size = 17, color = "#2A5783"),
        plot.caption = element_markdown(hjust = 0.5, color = "gray50", 
                                        margin = margin(t = 15)),
        plot.background = element_rect(fill = "white", color = NA)
  ) 


plot2 <- ggplot(merge3, aes(x = total_samples, y = se)) +
  geom_hex(bins = 30) +
  scale_fill_distiller(
    palette = "Spectral",
    direction = 1,
    name = "Count"
  ) +
  facet_wrap(~ homogeneous,
             labeller = as_labeller(c(
               "TRUE"  = "Homogeneous",
               "FALSE" = "Heterogeneous"
             ))) +
  labs(
    x     = expression(paste("Total Samples per ", 3^{rd}, " administrative Unit")),
    y     = "Design-based Standard Error",
    title = ""
  ) +
  theme_bw(base_size = 14) +
  theme(
    
    text = element_text(family = "Arial Narrow", color = "#2a2a2a"),
        plot.title = element_markdown(hjust = 0.5, lineheight = 1.2, 
                                      margin = margin(t = 10, b = 10)),
        plot.subtitle = element_markdown(hjust = 0.5, size = 16,
                                         color = "gray30", margin = margin(b = 20)),
        axis.title = element_text(face = "bold", size = 16),
        legend.position = "bottom",
        legend.text = element_text(size = 13),
        axis.text.x = element_text(angle = 90, hjust = 1, color = "gray40"),
        axis.text.y = element_text(color = "gray40"),
        # panel.grid.major.x = element_blank(),
        # panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        strip.background = element_rect(fill = "#F7F7F7", color = NA),
        strip.text = element_text(face="bold", size = 17, color = "#2A5783"),
        plot.caption = element_markdown(hjust = 0.5, color = "gray50", 
                                        margin = margin(t = 15)),
        plot.background = element_rect(fill = "white", color = NA)
  ) 

final_plot <- plot1 + plot2 + plot_layout(ncol = 2)
print(final_plot)
```



```{r}
design_based_estimates_admin1

data_ethiopia_1_filter <- data.frame(
  ADM1_EN = data_ethiopia_1$ADM1_EN,
  ADM2_EN = data_ethiopia_1$ADM2_EN,
  ADM3_EN = data_ethiopia_1$ADM3_EN
)

hhh <- merge(data_ethiopia_1_filter, design_based_estimates_admin2, by=c("ADM2_EN"))


hhhhh <- merge(hhh, design_based_estimates_admin1, by=c("ADM1_EN"))


```



```{r}

df_sd <- tibble(
  area           = seq_along(plot_plot_no$sd_beta_bino),
  `Without urban` = plot_plot_no$sd_beta_bino,
  `With urban`    = plot_plot$sd_beta_bino
) %>%
  pivot_longer(-area, names_to = "Model", values_to = "SD")
mean_df <- df_sd %>%
  group_by(Model) %>%
  summarize(mean_SD = mean(SD), .groups = "drop")
cbPalette <- c("Without urban" = "#0072B2", "With urban" = "#D55E00")

ggplot(df_sd, aes(x = area, y = SD, color = Model)) +
  geom_point(size = 3, shape=21, alpha=0.4) +
  geom_hline(
    data = mean_df,
    aes(yintercept = mean_SD, color = Model),
    linetype = "dashed",
    size = 0.8
  ) +
  scale_color_manual(values = cbPalette) +
  scale_shape_manual(values = c(16, 17)) +
  labs(
    x     = expression(paste("Total Samples per ", 2^{rd}, " administrative Unit")),
    y     = "Standard errors",
    # title = "Comparison of Posterior standard errors with and without urban covariate",
    subtitle = "",
    color = NULL, shape = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position    = "top",
    panel.grid.major.x = element_blank()
  )

```

```{r}
ggsave("plot_new_admin2.png", 
       width = 12, height = 7, dpi = 400, bg = "white")
```




MBG and my estimte

```{r}
est_2_unit_level_urban <- 
  st_read("result/unit_level_estimate_admin2_urban.shp")

est_3_unit_level_urban_mbg <- 
  read_csv("result/agg_summaries_all.csv")

est_3_unit_level_urban_mbg <- est_3_unit_level_urban_mbg %>% filter(
  vaccine == "mcv1" & year == 2019
)

names(est_3_unit_level_urban)

merge_both <- left_join(est_3_unit_level_urban, est_3_unit_level_urban_mbg, by=c("ADM3_EN_x"="ADM3_EN"))

plot(merge_both$mean, merge_both$md_bt_b)
merge_both$se_one <- merge_both$uppr_b_ - merge_both$lwr_bt_

p <- ggplot(merge_both, aes(x = md_bt_b, y = mean*100)) +

    geom_point(aes(size=se_one),color="#D35400", shape=21, alpha = 0.85, show.legend = TRUE) +
  
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "#3a86ff", linewidth = 1) +
  scale_color_viridis(
    option = "plasma", direction = -1, name = "Prediction SE"
  ) +
  # scale_size_continuous(name = "Prediction SE", range = c(2, 8)) +
  labs(
    x = "Model-based small area estimation \n (with urban variable)",
    y = "Model Based Geostatistics \n (MBG)",
    title = "",
    subtitle = "",
    caption = ""
  ) +
  stat_cor(method = "pearson", label.x.npc = 0.05, label.y.npc = 0.95,
           aes(label = ..r.label..), color = "#34495E", size=13) +
  theme_pubclean()+
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "grey30",
                                margin = margin(b = 15)),
    axis.title = element_text(face = "bold", size = 17),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text = element_text(size = 14, color = "grey40"),
    axis.ticks = element_line(color = "grey60", linewidth = 0.6),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "none"
  ) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf,
           color = "grey40", linewidth = 0.8) +
  annotate("segment", y = -Inf, yend = Inf, x = -Inf, xend = -Inf,
           color = "grey40", linewidth = 0.8)

print(p)

```

```{r}
ggsave("final_mbg_sae_admin2_se.png",
       width = 8, height = 5, dpi = 400, bg = "white")
```


```{r}
z <- qnorm(0.975)  # 1.96
se1 <- (merge_both$uppr_b_   - merge_both$lwr_bt_) / (2 * z)
se2 <- (merge_both$upper     - merge_both$lower  ) / (2 * z)
merge_both$se1 <- se1
merge_both$se2 <- se2
p <- ggplot(merge_both, aes(x = se1*100, y = se2*100)) +

    geom_point(color="#D35400", shape=21, alpha = 0.85, show.legend = TRUE) +
  
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "#3a86ff", 
              linewidth = 1) +
  scale_color_viridis(
    option = "plasma", direction = -1, name = "Prediction SE"
  ) +
  # scale_size_continuous(name = "Prediction SE", range = c(2, 8)) +
  labs(
    x = "Standard Error \n Model-based small area estimation \n (with urban variable)",
    y = "Standard Error \n Model Based Geostatistics \n (MBG)",
    title = "",
    subtitle = "",
    caption = ""
  ) +
  stat_cor(method = "pearson", label.x.npc = 0.05, label.y.npc = 0.95,
           aes(label = ..r.label..), color = "#34495E", size=13) +
  theme_pubclean()+
  theme(
    plot.title = element_text(face = "bold", size = 18, hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "grey30",
                                margin = margin(b = 15)),
    axis.title = element_text(face = "bold", size = 17),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text = element_text(size = 14, color = "grey40"),
    axis.ticks = element_line(color = "grey60", linewidth = 0.6),
    panel.background = element_blank(),
    panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "none"
  ) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf,
           color = "grey40", linewidth = 0.8) +
  annotate("segment", y = -Inf, yend = Inf, x = -Inf, xend = -Inf,
           color = "grey40", linewidth = 0.8)

print(p)
```



```{r}


#estimate normal distribution parameters from mean and confidence interval
estimate_normal_params <- function(mean_val, lower_ci, upper_ci, confidence_level = 0.95) {
  ci_width <- upper_ci - lower_ci
  z_score <- qnorm(1 - (1 - confidence_level)/2)  # 1.96 for 95% CI
  standard_error <- ci_width / (2 * z_score)
  
  return(list(mean = mean_val, sd = standard_error))
}

calc_exceed_prob <- function(mean_val, lower_ci, upper_ci, threshold = 0.9) {
  if (is.na(mean_val) | is.na(lower_ci) | is.na(upper_ci)) {
    return(NA)
  }
  if (lower_ci >= threshold) {
    return(0.999)
  }
  if (upper_ci <= threshold) {
    return(0.001)
  }
  params <- estimate_normal_params(mean_val, lower_ci, upper_ci)
  prob <- 1 - pnorm(threshold, mean = params$mean, sd = params$sd)
  prob <- pmax(0, pmin(1, prob))
  return(prob)
}
analyze_mcv1_coverage <- function(data, threshold = 0.9) {
  
  results <- data %>%
    mutate(
      prob_exceed = map_dbl(1:n(), ~calc_exceed_prob(
        estimate[.x], 
        lwr_bt_[.x], 
        uppr_b_[.x], 
        threshold
      )),
      #classify based on probability
      classification = case_when(
        prob_exceed >= 0.8 ~ "Very Likely (80%)",
        prob_exceed >= 0.5 ~ "Likely (50-80%)",
        prob_exceed >= 0.2 ~ "Possible (20-50%)",
        prob_exceed > 0 ~ "Unlikely (<20%)",
        TRUE ~ "Very Unlikely (0%)"
      )
    )
  
  return(results)
}
summarize_results <- function(results, threshold = 0.9) {
  summary_stats <- list(
    n_total = nrow(results),
    n_above_threshold = sum(results$prob_exceed > 0.5, na.rm = TRUE),
    mean_probability = mean(results$prob_exceed, na.rm = TRUE),
    median_probability = median(results$prob_exceed, na.rm = TRUE),
    overall_likelihood = mean(results$prob_exceed, na.rm = TRUE),
    high_confidence_count = sum(results$prob_exceed >= 0.8, na.rm = TRUE),
    classification_summary = table(results$classification)
  )
  
  return(summary_stats)
}
plot_probability_distribution <- function(results, threshold = 0.9) {
  p1 <- ggplot(results, aes(x = prob_exceed)) +
    geom_histogram(bins = 20, fill = "lightblue", color = "black", alpha = 0.7) +
    geom_vline(xintercept = 0.5, linetype = "dashed", color = "red", size = 1) +
    labs(
      title = paste("Distribution of Probabilities of Exceeding", threshold),
      subtitle = "Based on Normal Distribution Model",
      x = "Probability of Exceeding Threshold",
      y = "Count",
      caption = "Red line indicates 50% probability"
    ) +
    theme_minimal()
  p2 <- ggplot(results, aes(x = reorder(ADM3_EN_y, estimate))) +
    geom_point(aes(y = estimate, color = prob_exceed), size = 2) +
    geom_errorbar(aes(ymin = lwr_bt_, ymax = uppr_b_, color = prob_exceed), 
                  width = 0.2, alpha = 0.7) +
    geom_hline(yintercept = threshold, linetype = "dashed", color = "red", size = 1) +
    scale_color_gradient2(
      low = "red", mid = "yellow", high = "green", midpoint = 0.5,
      name = "Prob. of\nExceeding\nThreshold"
    ) +
    labs(
      title = paste("MCV1 Coverage Estimates with Uncertainty (Normal Model)"),
      subtitle = paste("Threshold =", threshold),
      x = "Area",
      y = "Coverage Estimate"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  return(list(prob_hist = p1, coverage_plot = p2))
}
run_mcv1_analysis <- function(data, threshold = 0.9) {
  data <- data %>%
    rename(estimate = md_bt_b) %>%
    select(ADM3_EN_y, estimate, lwr_bt_, uppr_b_)
  results <- analyze_mcv1_coverage(data, threshold)
  summary_stats <- summarize_results(results, threshold)
  plots <- plot_probability_distribution(results, threshold)
  cat("=== MCV1 Coverage Analysis Summary ===\n")
  cat("Threshold:", threshold, "\n")
  cat("Total areas analyzed:", summary_stats$n_total, "\n")
  cat("Areas likely to exceed threshold (>50% prob):", summary_stats$n_above_threshold, "\n")
  cat("Overall average probability:", round(summary_stats$overall_likelihood, 3), "\n")
  cat("Median probability:", round(summary_stats$median_probability, 3), "\n")
  cat("High confidence areas (80% prob):", summary_stats$high_confidence_count, "\n\n")
  
  cat("Classification breakdown:\n")
  print(summary_stats$classification_summary)
  print(plots$prob_hist)
  print(plots$coverage_plot)
  
  return(list(
    results = results,
    summary = summary_stats,
    plots = plots
  ))
}
analysis_results <- run_mcv1_analysis(est_2_unit_level_urban , threshold = 0.9)

#maybe not used
ggsave("new_90.png", width = 14, height = 10, bg = "white")

```



```{r}
# Bottom 20% Classification Confidence Analysis
# How confident are we in identifying the true bottom 20% given uncertainty?
# Bottom 20% Classification Probability Analysis
# For each district: probability of being labeled as bottom 20% given uncertainty
simulate_bottom20_probabilities <- function(data, n_simulations = 10000) {
  data <- data %>%
    rename(estimate = md_bt_b) %>%
    select(ADM3_EN_y, estimate, lwr_bt_, uppr_b_) %>%
    filter(!is.na(estimate) & !is.na(lwr_bt_) & !is.na(uppr_b_)) %>%
    mutate(district_id = row_number())
  
  n_districts <- nrow(data)
  bottom20_count <- ceiling(n_districts * 0.2)
  
  cat("Analyzing", n_districts, "districts\n")
  cat("Bottom 20% =", bottom20_count, "districts\n")
  cat("Running", n_simulations, "simulations...\n\n")
  bottom20_matrix <- matrix(FALSE, nrow = n_simulations, ncol = n_districts)
  
  #simulations
  for (sim in 1:n_simulations) {
    simulated_values <- runif(n_districts, 
                             min = data$lwr_bt_, 
                             max = data$uppr_b_)
    ranks <- rank(simulated_values, ties.method = "random")
    bottom20_districts <- ranks <= bottom20_count
    bottom20_matrix[sim, ] <- bottom20_districts
  }
  district_probabilities <- colMeans(bottom20_matrix)
  results <- data %>%
    mutate(
      prob_labeled_bottom20 = district_probabilities,
      point_estimate_rank = rank(estimate),
      point_estimate_bottom20 = point_estimate_rank <= bottom20_count
    ) %>%
    arrange(desc(prob_labeled_bottom20))
  
  return(list(
    results = results,
    n_districts = n_districts,
    bottom20_count = bottom20_count,
    simulation_matrix = bottom20_matrix
  ))
}

# categorize districts by probability
categorize_districts <- function(results_data) {
  
  results_data %>%
    mutate(
      probability_category = case_when(
        prob_labeled_bottom20 >= 0.8 ~ "Very High (80%)",
        prob_labeled_bottom20 >= 0.6 ~ "High (60-80%)",
        prob_labeled_bottom20 >= 0.4 ~ "Moderate (40-60%)",
        prob_labeled_bottom20 >= 0.2 ~ "Low (20-40%)",
        prob_labeled_bottom20 >= 0.05 ~ "Very Low (5-20%)",
        TRUE ~ "Minimal (<5%)"
      ),
      
      classification_status = case_when(
        point_estimate_bottom20 & prob_labeled_bottom20 >= 0.6 ~ "Stable Bottom 20%",
        point_estimate_bottom20 & prob_labeled_bottom20 < 0.6 ~ "Uncertain Bottom 20%",
        !point_estimate_bottom20 & prob_labeled_bottom20 >= 0.4 ~ "Potential Bottom 20%",
        TRUE ~ "Unlikely Bottom 20%"
      )
    )
}
create_probability_visualizations <- function(categorized_data, bottom20_count) {
  p1 <- ggplot(categorized_data, aes(x = prob_labeled_bottom20)) +
    geom_histogram(bins = 50, fill = "lightblue", color = "black", alpha = 0.7) +
    geom_vline(xintercept = 0.2, color = "red", linetype = "dashed", size = 1) +
    geom_vline(xintercept = 0.8, color = "orange", linetype = "dashed", size = 1) +
    labs(
      title = "Distribution of Bottom 20% Classification Probabilities",
      subtitle = "Each district's probability of being labeled as bottom 20%",
      x = "Probability of Being Labeled Bottom 20%",
      y = "Number of Districts",
      caption = "Red line = 20%, Orange line = 80%"
    ) +
    theme_minimal()
  
  #ordered by probability
  p2 <- ggplot(categorized_data, aes(x = reorder(ADM3_EN_y, prob_labeled_bottom20), 
                                     y = prob_labeled_bottom20)) +
    geom_col(aes(fill = probability_category), alpha = 0.8) +
    geom_hline(yintercept = c(0.2, 0.4, 0.6, 0.8), linetype = "dashed", alpha = 0.5) +
    scale_fill_manual(values = c(
      "Very High (80%)" = "darkred",
      "High (60-80%)" = "red",
      "Moderate (40-60%)" = "orange",
      "Low (20-40%)" = "yellow",
      "Very Low (5-20%)" = "lightblue",
      "Minimal (<5%)" = "lightgray"
    )) +
    labs(
      title = "Individual District Classification Probabilities",
      x = "District (ordered by probability)",
      y = "Probability of Being Labeled Bottom 20%",
      fill = "Probability Category"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6))
  
  #point estimate vs simulation probability
  p3 <- ggplot(categorized_data, aes(x = estimate, y = prob_labeled_bottom20)) +
    geom_point(aes(color = classification_status), size = 2, alpha = 0.7) +
    geom_hline(yintercept = 0.2, linetype = "dashed", alpha = 0.5) +
    geom_vline(xintercept = quantile(categorized_data$estimate, 0.2), 
               linetype = "dashed", alpha = 0.5) +
    scale_color_manual(values = c(
      "Stable Bottom 20%" = "darkgreen",
      "Uncertain Bottom 20%" = "orange", 
      "Potential Bottom 20%" = "blue",
      "Unlikely Bottom 20%" = "gray"
    )) +
    labs(
      title = "Point Estimate vs Classification Probability",
      x = "Point Estimate (Coverage)",
      y = "Probability of Being Labeled Bottom 20%",
      color = "Classification Status"
    ) +
    theme_minimal()
  
  return(list(prob_hist = p1, district_bars = p2, scatter_plot = p3))
}

# analysis function
analyze_bottom20_classification_probability <- function(data, n_simulations = 10000) {
  
  cat("=== BOTTOM 20% CLASSIFICATION PROBABILITY ANALYSIS ===\n\n")
  
  #  simulation
  sim_results <- simulate_bottom20_probabilities(data, n_simulations)
  categorized_data <- categorize_districts(sim_results$results)
  cat("SIMULATION SUMMARY:\n")
  cat("Total districts:", sim_results$n_districts, "\n")
  cat("Bottom 20% size:", sim_results$bottom20_count, "districts\n\n")
  
  # category breakdown
  cat("PROBABILITY CATEGORY BREAKDOWN:\n")
  category_summary <- table(categorized_data$probability_category)
  print(category_summary)
  cat("\n")
  cat("CLASSIFICATION STATUS BREAKDOWN:\n")
  status_summary <- table(categorized_data$classification_status)
  print(status_summary)
  cat("\n")
  cat("KEY STATISTICS:\n")
  cat("Mean probability across all districts:", 
      round(mean(categorized_data$prob_labeled_bottom20), 3), "\n")
  cat("Districts with >80% probability:", 
      sum(categorized_data$prob_labeled_bottom20 >= 0.8), "\n")
  cat("Districts with >60% probability:", 
      sum(categorized_data$prob_labeled_bottom20 >= 0.6), "\n")
  cat("Districts with 20-80% probability (uncertain):", 
      sum(categorized_data$prob_labeled_bottom20 >= 0.2 & 
          categorized_data$prob_labeled_bottom20 < 0.8), "\n\n")
  cat("TOP 10 DISTRICTS BY CLASSIFICATION PROBABILITY:\n")
  top_districts <- head(categorized_data %>% 
                       select(ADM3_EN_y, estimate, prob_labeled_bottom20, 
                              probability_category, classification_status), 10)
  print(top_districts)
  cat("\n")
  
  #point-estimate bottom 20% with low simulation probability
  uncertain_bottom20 <- categorized_data %>%
    filter(point_estimate_bottom20 & prob_labeled_bottom20 < 0.6) %>%
    arrange(prob_labeled_bottom20) %>%
    select(ADM3_EN_y, estimate, prob_labeled_bottom20, classification_status)
  
  if(nrow(uncertain_bottom20) > 0) {
    cat("POINT-ESTIMATE BOTTOM 20% WITH LOW SIMULATION PROBABILITY:\n")
    print(uncertain_bottom20)
    cat("\n")
  }
  
  # non-point-estimate districts with high simulation probability
  potential_bottom20 <- categorized_data %>%
    filter(!point_estimate_bottom20 & prob_labeled_bottom20 >= 0.4) %>%
    arrange(desc(prob_labeled_bottom20)) %>%
    select(ADM3_EN_y, estimate, prob_labeled_bottom20, classification_status)
  
  if(nrow(potential_bottom20) > 0) {
    cat("NON-POINT-ESTIMATE DISTRICTS WITH HIGH SIMULATION PROBABILITY:\n")
    print(potential_bottom20)
    cat("\n")
  }
  plots <- create_probability_visualizations(categorized_data, sim_results$bottom20_count)
  print(plots$prob_hist)
  print(plots$district_bars)
  print(plots$scatter_plot)
  stable_count <- sum(categorized_data$classification_status == "Stable Bottom 20%")
  uncertain_count <- sum(categorized_data$classification_status == "Uncertain Bottom 20%")
  
  cat("=== INTERPRETATION SUMMARY ===\n")
  cat("Out of", sim_results$bottom20_count, "districts in point-estimate bottom 20%:\n")
  cat("- Stable (60% simulation probability):", stable_count, "\n")
  cat("- Uncertain (<60% simulation probability):", uncertain_count, "\n")
  cat("Classification certainty:", round(stable_count/sim_results$bottom20_count*100, 1), "%\n")
  
  return(list(
    simulation_results = sim_results,
    categorized_data = categorized_data,
    plots = plots,
    summary_stats = list(
      stable_districts = stable_count,
      uncertain_districts = uncertain_count,
      classification_certainty = stable_count/sim_results$bottom20_count
    )
  ))
}

# Run:
results <- analyze_bottom20_classification_probability(est_2_unit_level_urban, n_simulations = 10000)
```

```{r}

est_3_unit_level_urban <- 
  st_read("result/unit_level_estimate_admin2_urban_new_new.shp")

## ---------------------------------------------------------------
##  1.  Add P(coverage > 0.90) for every ADM2
## ---------------------------------------------------------------

# est_3_unit_level_urban <- st_drop_geometry(est_3_unit_level_urban)
est_3_unit_level_urban$md_bt_b <- est_3_unit_level_urban$md_bt_b/100
est_3_unit_level_urban$sd_bt_b <- est_3_unit_level_urban$sd_bt_b/100

df_prob <- est_3_unit_level_urban %>% 
  rowwise() %>%                             # one row at a time
  mutate(
    p_gt90 = 1 - ptnorm(
      0.9,
      mean  = md_bt_b,
      sd    = sd_bt_b
      # lower = 0, upper = 1                 # truncate to [0,1]
    )
  ) %>% 
  ungroup()


## ---------------------------------------------------------------
##  2.  Overall summaries 
## ---------------------------------------------------------------
overall_equal_weight  <- mean(df_prob$p_gt90, na.rm = TRUE)

share_high_confidence <- mean(df_prob$p_gt90 >= 0.80, na.rm = TRUE)   
prob_all_hit_target   <- prod(df_prob$p_gt90,   na.rm = TRUE)         

```



```{r}
df   <- est_3_unit_level_urban %>%
    mutate(
        sd_use = ifelse(!is.na(sd_bt_b), sd_bt_b,
                        (upprb - lwrbt) / (2 * qnorm(.975)))
    )
df <- df %>%
    mutate(
        n_eff = pmax((md_bt_b * (1 - md_bt_b)) / sd_use^2 - 1, 0.01)
    )
df <- df %>%
    mutate(
        alpha = 1 + md_bt_b * n_eff,
        beta  = 1 + (1 - md_bt_b) * n_eff
    )
df_prob_beta <- df %>%
    mutate(
        p_gt90 = 1 - pbeta(0.90, alpha, beta)
    )
overall_equal_weight  <- mean(df_prob_beta$p_gt90, na.rm = TRUE)
share_high_confidence <- mean(df_prob_beta$p_gt90 >= 0.80, na.rm = TRUE)
prob_all_hit_target   <- prod(df_prob_beta$p_gt90,   na.rm = TRUE)  # independence

cat("\n--- Posterior exceedance results -----------------------------------\n",
    sprintf("Mean   P(coverage > 90 %%):  %.3f", overall_equal_weight), "\n",
    sprintf("Share  with 80 %% chance  :  %.1f %%", 100*share_high_confidence), "\n",
    sprintf("Prob. all units > 90 %%    :  %.4f", prob_all_hit_target), "\n")
```


```{r}
 ggsave("map1_new_final.png", width = 16, height = 12, bg = "white")
```


```{r}

df <- est_3_unit_level_urban %>%
  mutate(
    sd_use = ifelse(!is.na(sd_bt_b),
                    sd_bt_b,
                    (upprb - lwrbt) / (2 * qnorm(.975)))
  )
df <- df %>%
  mutate(
    n_eff = pmax((md_bt_b * (1 - md_bt_b)) / sd_use^2 - 1, 0.01),
    alpha = 1 + md_bt_b * n_eff,
    beta  = 1 + (1 - md_bt_b) * n_eff
  )
set.seed(2025)
B  <- 2000
n  <- nrow(df)

draw_matrix <- replicate(
  B,
  rbeta(n, df$alpha, df$beta)   # one draw per unit
)

bottom_cut   <- floor(0.20 * n)          # bottomquintile size
in_bottom_MC <- apply(draw_matrix, 2, function(sim){
  idx               <- order(sim)       
  out               <- logical(n)
  out[idx[1:bottom_cut]] <- TRUE
  out
})
prob_bottom <- rowMeans(in_bottom_MC)
df_stab <- df %>%
  mutate(
    bottom_point = md_bt_b <= quantile(md_bt_b, 0.20, type = 8),
    prob_bottom  = prob_bottom,
    certainty_band = cut(
      prob_bottom,
      breaks = c(0, .30, .40, .50, .60, .70, .80, 1),
      labels = c("<30%", "3040%", "4050%", "5060%",
                 "6070%", "7080%", "80%"),
      right = FALSE
    )
  )
stab_summary_tbl <- df_stab %>%
  filter(bottom_point) %>%
  group_by(certainty_band) %>%
  summarise(
    n_units = n(),
    share   = scales::percent(n_units / n(), accuracy = 0.1),
    .groups = "drop"
  )

print(stab_summary_tbl)


```


```{r}

df_map <- df_stab %>%
  mutate(
    highlight = prob_bottom >= 0.70,
    fill_col  = if_else(highlight, "highlight", "other")
  )
centroids <- st_centroid(df_map)                       # still sf
centroids <- cbind(centroids, st_coordinates(centroids))  
font_add_google("Playfair Display", "playfair")  
font_add_google("Source Sans Pro", "source")    
font_add_google("Inter", "inter")               
showtext_auto()
create_award_winning_map <- function(df_map, centroids) {
  n_highlighted <- sum(df_map$highlight, na.rm = TRUE)
  total_units <- nrow(df_map)
  pct_highlighted <- round((n_highlighted / total_units) * 100, 1)
  region_stats <- df_map %>%
    st_drop_geometry() %>%
    group_by(ADM1_EN) %>%
    summarise(
      total_districts = n(),
      high_risk_districts = sum(highlight, na.rm = TRUE),
      pct_at_risk = round((high_risk_districts / total_districts) * 100, 1),
      .groups = 'drop'
    ) %>%
    arrange(desc(pct_at_risk))
  top_regions <- region_stats %>%
    filter(high_risk_districts > 0) %>%
    slice_head(n = 4) %>%
    pull(ADM1_EN)
  colors_palette <- list(
    background = "#FFFFFF",
    water = "#E3F2FD",
    other = "#F5F5F5",
    highlight = "#1565C0",
    highlight_dark = "#0D47A1",
    text_primary = "#1A1A1A",
    text_secondary = "#666666",
    accent = "#FF6B35",
    border = "grey"
  )
  p <- ggplot() +
    theme(
      plot.background = element_rect(fill = colors_palette$background, color = NA),
      panel.background = element_rect(fill = colors_palette$background, color = NA)
    ) +
    geom_sf(
      data = df_map,
      aes(fill = fill_col),
      color = colors_palette$border,
      size = 0.25,
      alpha = 0.9
    ) +
    geom_sf(
      data = subset(df_map, highlight == TRUE),
      fill = colors_palette$highlight_dark,
      color = "transparent",
      alpha = 0.15,
      position = position_nudge(x = 0.001, y = -0.001)
    ) +
    scale_fill_manual(
      values = c(
        other = colors_palette$other, 
        highlight = colors_palette$highlight
      ),
      guide = "none"
    ) +
    geom_text_repel(
      data = subset(centroids, highlight),
      aes(x = X, y = Y, label = paste0(ADM2_EN_y, "\n(", ADM1_EN, ")")),
      family = "source",
      size = 3,
      fontface = "bold",
      color = colors_palette$text_primary,
      bg.color = alpha("white", 0.95),
      bg.r = 0.12,
      box.padding = 0.5,
      point.padding = 0.4,
      segment.color = colors_palette$highlight,
      segment.alpha = 0.7,
      segment.size = 0.8,
      # force = 4,
      max.overlaps = Inf,
      min.segment.length = 0.2,
      seed = 42,  # For reproducible label placement
      lineheight = 0.9
    ) +
    coord_sf(
      expand = FALSE,
      clip = "off"
    ) +
    labs(
      title = "CRITICAL VACCINATION GAPS BY REGION",
      subtitle = paste0("District-level MCV1 underperformance risk across administrative regions  ",
                       n_highlighted, " districts in ", length(top_regions), " regions show 70% risk  ",
                       "Most affected: ", paste(head(top_regions, 2), collapse = ", ")),
      caption = paste0("Methodology: Bayesian posterior resampling of administrative vaccination data\n",
                      "Labels show District (Region)  Districts requiring immediate intervention focus\n",
                      "Regional breakdown: ", nrow(region_stats), " regions analyzed")
    ) +
    theme_void(base_family = "source") +
    theme(
      plot.title = element_text(
        family = "playfair",
        face = "bold",
        size = 28,
        hjust = 0.5,
        color = colors_palette$text_primary,
        margin = margin(b = 8, t = 25),
        lineheight = 1.1
      ),
      
      plot.subtitle = element_text(
        family = "inter",
        size = 13,
        hjust = 0.5,
        color = colors_palette$text_secondary,
        margin = margin(b = 30),
        lineheight = 1.3
      ),
      
      plot.caption = element_text(
        family = "inter",
        size = 10,
        hjust = 0,
        color = colors_palette$text_secondary,
        margin = margin(t = 25, l = 20),
        lineheight = 1.2
      ),
      plot.margin = margin(40, 50, 40, 50, "pt"),
      panel.grid = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      plot.background = element_rect(
        fill = colors_palette$background, 
        color = NA
      ),
      panel.background = element_rect(
        fill = colors_palette$background, 
        color = NA
      )
    )
  
  return(p)
}
create_artistic_masterpiece <- function(df_map, centroids) {
  region_stats <- df_map %>%
    st_drop_geometry() %>%
    group_by(ADM1_EN) %>%
    summarise(
      total_districts = n(),
      high_risk_districts = sum(highlight, na.rm = TRUE),
      pct_at_risk = round((high_risk_districts / total_districts) * 100, 1),
      .groups = 'drop'
    ) %>%
    arrange(desc(pct_at_risk))
  colors_palette <- list(
    background = "#F8F9FA",
    water = "#E3F2FD",
    other = "#F5F5F5",
    highlight = "#1565C0",
    highlight_dark = "#0D47A1",
    text_primary = "#1A1A1A",
    text_secondary = "#666666",
    accent = "#FF6B35",
    border = "white"
  )
  base_map <- create_award_winning_map(df_map, centroids)
  enhanced_map <- base_map +
    annotate(
      "text",
      x = Inf, y = Inf,
      label = "N ",
      hjust = 1.2, vjust = 1.2,
      family = "inter",
      size = 4,
      color = "#999999",
      fontface = "bold"
    ) +
    annotate(
      "segment",
      x = -Inf, xend = -Inf,
      y = -Inf, yend = -Inf,
      color = "#999999",
      size = 0.8,
      alpha = 0.7
    ) +
    annotate(
      "label",
      x = Inf, y = -Inf,
      label = paste0("REGIONAL RISK SUMMARY\n",
                    paste0(region_stats$ADM1_EN[1:min(3, nrow(region_stats))], 
                          ": ", region_stats$high_risk_districts[1:min(3, nrow(region_stats))], 
                          " zones", collapse = "\n")),
      hjust = 1.1, vjust = -0.1,
      family = "inter",
      fontface = "bold",
      size = 3.2,
      color = colors_palette$highlight,
      fill = alpha("white", 0.95),
      label.padding = unit(0.5, "lines"),
      label.r = unit(0.25, "lines")
    )
  
  return(enhanced_map)
}
export_map <- function(map_plot, filename = "mcv1_risk_map", 
                      width = 16, height = 12, dpi = 300) {
  
  ggsave(
    filename = paste0(filename, ".png"),
    plot = map_plot,
    width = width,
    height = height,
    dpi = dpi,
    bg = "#FFFFFF",
    device = "png",
    type = "cairo"
  )
  ggsave(
    filename = paste0(filename, ".pdf"),
    plot = map_plot,
    width = width,
    height = height,
    device = "pdf",
    bg = "#FFFFFF"
  )
  
  cat("Maps exported as:", paste0(filename, c(".png", ".pdf")), "\n")
}

artistic_map <- create_artistic_masterpiece(df_map, centroids)
print(artistic_map)
export_map(artistic_map, 'mcv1_vaccination_risk_admin_2')

```

